---
title: "Myrtle Rust Risk Assessment"
format: html
editor: visual
---

```{r}
library(data.table)
library(sf)
library(tidyverse)
library(raster)
library(terra)
library(ggspatial)
library(ggtext)
library(patchwork)
library(classInt)
library(RColorBrewer)
library(gt)
library(webshot2)
library(flextable)
library(officer)
library(viridisLite)
library(rnaturalearth)
library(rnaturalearthdata)
library(scales)
library(prioritizr)
library(gurobi)
```

# Susceptibility Map

This portion of the work adapts <https://github.com/meryltheng/carbon-contribution.git> to produce a national "susceptibility map" by approximating the host density for the Australian National Vegetation Information System's (NVIS) Major Vegetation Groups (MVG) and Major Vegetation Subgroups (MVS). To produce such approximations required the estimation of proportional impact in the form of lost/dead trees per hectare. This was done using TK Le's approximations of genus level impacts from myrtle rust which gave an impact score between 0 and 1. This can be thought of as proportional damage. These estimations where then combined with stand-level tree and plant data originally collected and summarised by Meryl Theng to assess impacts to biomass. We then adapted to assess only tree information for a given plot rather than allometric data

## Data preparation

The following is adapted from 0_Data prep.R code written by Meryl Theng <https://github.com/meryltheng/carbon-contribution.git>.

Code remains mostly unchanged with only some renaming.

### Read NVIS data (MVG and MVS)

```{r}
"----------- Read NVIS data (MVG and MVS) ---------------"
# NVIS extant data. Available at: https://www.dcceew.gov.au/environment/environment-information-australia/national-vegetation-information-system/data-products
# read NVIS raster layers (already projected to Australian Albers)
# read MVG layer
raster_file = 'Data/Spatial/NVIS_MVG_RAS_AA.tif'
MVG <- raster(raster_file)

# read MVS layer
raster_file = 'Data/Spatial/NVIS_MVS_RAS_AA.tif'
MVS <- raster(raster_file)

# Define commonly required projections (for crs transformation later)
LatLong <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
Albers  <- '+proj=aea +lat_1=-18 +lat_2=-36 +lat_0=0 +lon_0=132 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs'
```

### Biomass Plot Library data

```{r}
# https://portal.tern.org.au/metadata/TERN/fc4a7249-ebb2-4ada-8e06-b552bfb297a3
# please download the necessary data file 'biolib_treelist.csv' from the above 
# link and place it in the Data/Biomass Plot Library/ folder (>100MB, so not included in the repo)

# read and subset to 'recent' data (2013 onwards) 
file_name = 'Data/Biomass Plot Library/biolib_treelist.csv' # tree-level data
all_data <- fread(file_name)
bpl_data <- all_data[obs_time >= "2013-01-01"]
length(unique(bpl_data$tree)) # all records of unique trees (might wanna double-check)
rm(all_data); gc()

# add genus (other cols are non-impt placeholders)
bpl_data[, c("genus", "sp", "spp", "blah") := tstrsplit(species, " ", fixed=TRUE)]
bpl_data$family <- NA

# match record locations to MVG/MVS class
bpl_data_sf <- sf::st_as_sf(bpl_data, coords = c("longitude", "latitude"), crs = LatLong) # GDA94
# reproject to Australian Albers
bpl_data_AA <- sf::st_transform(bpl_data_sf,
                                 crs = Albers)

# append MVG to site info
mvgValue = extract(MVG, sf::st_coordinates(bpl_data_AA))
bpl_data_AA <- cbind(bpl_data_AA, mvgValue)

# append MVS to site info
mvsValue = extract(MVS, sf::st_coordinates(bpl_data_AA))
bpl_data_AA <- cbind(bpl_data_AA, mvsValue)

# reformat to standardise
bpl_data_input <- bpl_data_AA |>
  dplyr::select(geometry = geometry,
         #longitude = longitude,
         #latitude = latitude,
         site = site,
         sitearea_ha = sitearea_ha, # total area of site (sum of all subplots)
         obs_time = obs_time,
         plantID = tree,
         family = family,
         genus = genus,
         species = species,
         measurement = measurement, # method used to measure diameter
         diameter = diameter,
         height = ht, # tree height; not alws avail
         agb_est = agb_drymass, # estimated; we won't use this, will use our own calculations for standardisation (note: BPL estimates look to be in the wrong scale not in kg)
         mvgValue = mvgValue,
         mvsValue = mvsValue
         )

# save(bpl_data_AA, file = "bpl_data_AA.RData")
```

### Robson Creek Rainforest data

```{r}
# https://portal.tern.org.au/metadata/TERN/5a864ec1-f780-4104-a242-7ff00e1f4962

# read and subset to 'recent' data (2013 onwards) 
file_name = 'Data/Robson Creek Rainforest/Robson_Creek_diameter_height_biomass_data.csv' # tree-level data
all_data <- fread(file_name)
robcreek_data <- all_data |>
  mutate(obs_time = as.Date(phenomenonTime, "%d/%m/%Y")) |>
  filter(obs_time >= "2013-01-01") |>
  arrange(obs_time) |> 
  group_by(plantId) |>  # one plant has multuple observations
  slice(n()) # last row (most recent observation of the plant)
rm(all_data); gc()

# match record locations to MVG/MVS class
robcreek_data_sf <- sf::st_as_sf(robcreek_data, coords = c("longitude", "latitude"), crs = LatLong) # GDA94
# reproject to Australian Albers
robcreek_data_AA <- sf::st_transform(robcreek_data_sf,
                                crs = Albers)

# append MVG to site info
mvgValue = extract(MVG, sf::st_coordinates(robcreek_data_AA))
robcreek_data_AA <- cbind(robcreek_data_AA, mvgValue)

# append MVS to site info
mvsValue = extract(MVS, sf::st_coordinates(robcreek_data_AA))
robcreek_data_AA <- cbind(robcreek_data_AA, mvsValue)

# reformat to standardise
robcreek_data_input <- robcreek_data_AA |>
  dplyr::select(geometry = geometry,
         #longitude = longitude,
         #latitude = latitude,
         site = siteId,
         sitearea_ha = plotLength_metres, 
         obs_time = obs_time,
         plantID = plantId,
         family = family,
         genus = genus,
         species = scientificName,
         measurement = stemDiameterPointOfMeasurement_metres, # method used to measure diameter
         diameter = stemDiameter_centimetres,
         height = stemHeight_metres, # tree height; not alws avail
         agb_est = aboveGroundBiomass_kilograms, # estimated; we won't use this, will use our own calculations for standardisation
         mvgValue = mvgValue,
         mvsValue = mvsValue
  ) |>
  mutate(site = rep("Robertson Creek", nrow(robcreek_data_AA)), 
         sitearea_ha = rep(1, nrow(robcreek_data_AA)), # 1ha
         measurement = rep("D130", nrow(robcreek_data_AA)), # diameter at 130cm
         )

```

### Calperum Mallee data

```{r}
# https://portal.tern.org.au/metadata/TERN/c8d054b2-9dfc-4328-8251-f0fd9276083d

# read and subset to 'recent' data (2013 onwards) 
file_name = 'Data/Calperum Mallee/Calperum_Mallee_diameter_height_biomass_data.csv' # tree-level data
all_data <- fread(file_name)
calperum_data <- all_data |>
  mutate(obs_time = as.Date(phenomenonTime, "%d/%m/%Y")) |>
  filter(obs_time >= "2013-01-01") |>
  arrange(obs_time) |> 
  group_by(plantId) |>  # one plant has multuple observations
  slice(n()) # last row (most recent observation of the plant); lol 85 plants
rm(all_data); gc()

# match record locations to MVG/MVS class
calperum_data_sf <- sf::st_as_sf(calperum_data, coords = c("longitude", "latitude"), crs = LatLong) # GDA94
# reproject to Australian Albers
calperum_data_AA <- sf::st_transform(calperum_data_sf,
                                     crs = Albers)

# append MVG to site info
mvgValue = extract(MVG, sf::st_coordinates(calperum_data_AA))
calperum_data_AA <- cbind(calperum_data_AA, mvgValue)

# append MVS to site info
mvsValue = extract(MVS, sf::st_coordinates(calperum_data_AA))
calperum_data_AA <- cbind(calperum_data_AA, mvsValue)

# reformat to standardise
calperum_data_input <- calperum_data_AA |>
  dplyr::select(geometry = geometry,
         #longitude = longitude,
         #latitude = latitude,
         site = siteId,
         sitearea_ha = plotLength_metres, 
         obs_time = obs_time,
         plantID = plantId,
         family = family,
         genus = genus,
         species = scientificName,
         measurement = stemDiameterPointOfMeasurement_metres, # method used to measure diameter
         diameter = stemDiameter_centimetres,
         height = stemHeight_metres, # tree height; not alws avail
         agb_est = aboveGroundBiomass_kilograms, # estimated; we won't use this, will use our own calculations for standardisation
         mvgValue = mvgValue,
         mvsValue = mvsValue
  ) |>
  mutate(site = rep("Calperum Mallee", nrow(calperum_data_AA)), 
         sitearea_ha = rep(1, nrow(calperum_data_AA)), # 1ha
         measurement = rep("D10", nrow(calperum_data_AA)), # diameter at 10cm
  )
```

### Compile and save all data

```{r}
save(bpl_data_input,
     robcreek_data_input,
     calperum_data_input,
     file = "Input/data_clean.RData")
```

## Tree data summarisations

The following is adapted 1_AGB contribution.R code written by Meryl Theng <https://github.com/meryltheng/carbon-contribution.git>.

Code has been mostly streamlined to remove unnecessary calculations and data, primarily calculations of biomass. Added additional summarisations by MVS in addition to MVG.

```{r}
# load data in
load("Input/data_clean.RData")
# merge standardised datasets
all_data <- rbind(bpl_data_input,
                  calperum_data_input,
                  robcreek_data_input)

# remove records without genus/species information
all_data <- all_data |>
  filter(!is.na(genus), species != "Tree") 

# add sizeclass as a column where DBH is grouped 0-30, 31-60, 61+
all_data$sizeclass <- cut(all_data$diameter,
                         breaks = c(0, 30, 60, Inf),
                         labels = c("0-30", "31-60", "61+"),
                         right = FALSE)
# Consolidate genera misspellings
all_data$genus<-gsub("Grevillia", "Grevillea", all_data$genus)

# Grouping all data by genus, sizeclass
all_data_std <- all_data |>
  group_by(site, genus, sizeclass) |>
  summarise(geometry = unique(geometry),
            site = unique(site),
            sitearea_ha = unique(sitearea_ha),
            genus = unique(genus),
            mvgValue = unique(mvgValue),
            mvsValue = unique(mvsValue),
            trees = n(),
            trees_per_ha = n()/unique(sitearea_ha)) # standardised to per ha

# Grouping by mvg
all_data_mvg <- all_data_std |>
  group_by(mvgValue) |>
  mutate(n_sites_per_mvg = length(unique(site))) |>
  group_by(mvgValue, genus, sizeclass) |>
  summarise(n_sites_per_mvg = unique(n_sites_per_mvg),
            genus = unique(genus),
            mvgValue = unique(mvgValue),
            trees_per_ha = mean(trees_per_ha, na.rm = T)) # standardised to per ha

# grouping by mvs
all_data_mvs <- all_data_std |>
  group_by(mvsValue) |>
  mutate(n_sites_per_mvs = length(unique(site))) |>
  group_by(mvsValue, genus, sizeclass) |>
  summarise(n_sites_per_mvs = unique(n_sites_per_mvs),
            genus = unique(genus),
            mvsValue = unique(mvsValue),
            trees_per_ha = mean(trees_per_ha, na.rm = T)) # standardised to per ha

```

## Damage calculations

THe following part of the script is a modified version of the damages.R source file(?)/ function(?) which was originally to be passed into the 6_Damages plus.R script. However, this is no longer necessary as we are not iterating over climate scenarios in the current work. The following code chunk takes the damage samples produced by TK's estimations and uses them to calculate firs the impact to plants based on their genus. This is as simple as averaging within each sample. This is followed by calculations of the impact to the broader MVG. This is done by taking the tree library information and calculating first the how many trees/plants will be impacted per hectare and then grouping by the MVG. Once this is a complete, 1000 samples of damage are produced. This is then repeated for MVS. These 1000 samples are then averaged to produce a single dataframe with values then used to produce the map

### Genus level

```{r}
impact_dir = 'Python/MR_samples'

# Read in & summarise each sample at genus level
damage_samples_genus <- list.files(path = impact_dir, 
                                   pattern = "*.csv", 
                                   full.names = TRUE) |>
  map(function(x){
    read.csv(x) |>
      mutate(Genus = tools::toTitleCase(Genus)) |>  # <-- standardise to first letter caps
      group_by(Genus) |>
      summarise(Impact = mean(Damage, na.rm = TRUE), .groups = "drop") |>
      filter(!is.na(Impact)) |>
      rename(genus = Genus)
  })

# Stack all 1000 runs and summarise across them 
genus_impacts <- damage_samples_genus |>
  bind_rows() |>
  group_by(genus) |>
  summarise(
    Impact_median = round((100*median(Impact, na.rm = TRUE)), digits = 1),
    Impact_mean  = round((100*mean(Impact, na.rm = TRUE)), digits = 1),
    Impact_q025  = quantile(Impact, 0.025, na.rm = TRUE),
    Impact_q975  = quantile(Impact, 0.975, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    # combine CI into one column as [x%, y%]
    CI95 = paste0(
      "[",
      scales::percent(Impact_q025, accuracy = 0.1), ", ",
      scales::percent(Impact_q975, accuracy = 0.1),
      "]"
    )
  ) |>
  left_join(all_data_std |>
              group_by(genus) |>
              summarise(`In tree data` = TRUE, .groups = "drop"),
            by = "genus") |>
  mutate(`In tree data` = if_else(is.na(`In tree data`), FALSE, `In tree data`)) |>
  dplyr::select(genus, Impact_mean, Impact_median, CI95, `In tree data`)



# Create thesis-ready flextable
ft_genus_impacts <- genus_impacts |>
  mutate(`In tree data` = if_else(`In tree data`, "Yes", "No")) |>
  arrange(desc(Impact_median)) |>
  dplyr::select(genus, Impact_median, CI95, `In tree data`) |>
  flextable() |>
  
  # Header labels
  set_header_labels(
    genus = "Genus",
    Impact_median = "Median impact (%)",
    CI95 = "95% credible interval",
    `In tree data` = "In tree data?"
  ) |>
  
  # Header styling
  bold(part = "header") |>
  bg(part = "header", bg = "#D3D3D3") |>   # light gray header
  color(part = "header", color = "black") |>
  
  # Zebra striping for rows
  theme_zebra() |>
  
  # Align columns
  align(j = "Impact_median", align = "center", part = "header") |>
  align(j = "In tree data", align = "center", part = "header") |>
  align(j = "CI95", align = "center", part = "header") |>
  align(j = "Impact_median", align = "center") |>
  align(j = "CI95", align = "center") |>
  align(j = "In tree data", align = "center") |>
  
  # Italicize genus column
  italic(j = "genus") |>
  
  # Adjust font size and autofit
  fontsize(size = 10, part = "all") |>
  width(j = 3, width = 1.3, unit = "in") |>
  
  # Add black horizontal lines
  hline_top(part = "header", border = fp_border(color="black", width = 2)) |>
  hline_bottom(part = "header", border = fp_border(color="black", width = 2)) |>
  hline(i = 1:nrow(genus_impacts), border = fp_border(color="black", width = 1))

# Preview in R
ft_genus_impacts

#Save as image
# Save flextable as PNG
save_as_image(ft_genus_impacts, path = "Output/images/genus_impacts_final.png", zoom = 2)


# Export to Word
read_docx() |>
  body_add_par("Genus-level Impacts", style = "heading 1") |>
  body_add_flextable(ft_genus_impacts) |>
  print(target = "Output/documents/thesis_genus_impacts_table.docx")
```

### MVG damage calculations

```{r}
tree_library = all_data_mvg
impact_dir = 'Python/MR_samples'
# sampling damage to each genus within each mvg
damage_samples_mvg <- list.files(path = impact_dir, pattern = "*.csv", full.names = T) |>
    map(function(x){
      # aggregate species impacts by genus
      impact_by_genus <- read.csv(x) |>
        group_by(Genus) |>
        summarise(Impact = mean(Damage, na.rm = T)) |>
        filter(!is.na(Impact)) |>
        rename(genus = Genus)
      # calculate impact to individual trees by MVG
      impact_mvg <- left_join(tree_library, impact_by_genus, by = c('genus')) |> # Assign based on genus match 
        mutate(Impact = if_else(is.na(Impact), 0, Impact, NA)) |> # if no info, impact = 0
        mutate(Impact_trees = trees_per_ha * Impact) |> # calc absolute impact to trees per ha
        group_by(mvgValue) |>
        # summarise(Impact_trees = sum(Impact_trees),
        #           Impact_prop = sum(Impact_trees) / sum(trees_per_ha)) |>
        st_drop_geometry() |>
        rename(MVG = mvgValue) |>
        ungroup() |>
        complete(MVG = c(1:32,99), fill = list(Impact_trees = 0, Impact = 0, trees_per_ha = 0))
      return(impact_mvg)
    })



# Damage_samples is the existing list of 1000 tibbles
# Average across 1000 trials
damage_avg_mvg <- damage_samples_mvg |>
  bind_rows() |>          # stack 1000
  group_by(MVG, genus, sizeclass) |>              # keep sizeclass
  summarise(
    Impact_mean       = mean(Impact, na.rm = TRUE),
    Impact_q025       = quantile(Impact, 0.025, na.rm = TRUE),
    Impact_q975       = quantile(Impact, 0.975, na.rm = TRUE),
    
    ImpactTrees_mean  = mean(Impact_trees, na.rm = TRUE),
    ImpactTrees_q025  = quantile(Impact_trees, 0.025, na.rm = TRUE),
    ImpactTrees_q975  = quantile(Impact_trees, 0.975, na.rm = TRUE),
    
    # keep stable attributes from the first run (identical across runs)
    trees_per_ha      = first(trees_per_ha),
    n_sites_per_mvg   = first(n_sites_per_mvg),
    .groups = "drop"
  ) |>
  arrange(MVG, genus, sizeclass)

# Output to a csv
write.csv(damage_avg_mvg, "Output/impact_avg_by_MVG_genus_sizeclass.csv", )
damage_avg_mvg <- read_csv("Output/impact_avg_by_MVG_genus_sizeclass.csv")

# Extract information on to a summarised MVG df
susceptibility_mvg <- damage_avg_mvg |>
  group_by(MVG) |>
  summarise(
    ImpactTrees_tot   = sum(ImpactTrees_mean, na.rm = TRUE),
    trees_per_ha_tot  = sum(trees_per_ha, na.rm = TRUE),
    Impact_proportion = (ImpactTrees_tot / trees_per_ha_tot),
    n_sites_per_mvg   = first(n_sites_per_mvg),
    .groups = "drop"
  ) |>
  replace_na(list(Impact_proportion = 0, n_sites_per_mvg = 0)) |> #data availability flag
  mutate(has_tree_data = if_else(n_sites_per_mvg > 0, TRUE, FALSE))

# Save MVGs with no tree data for overlay/masking
mvgs_no_tree_data <- susceptibility_mvg |> 
  filter(has_tree_data == FALSE) |>
  dplyr::select(MVG)

```

#### MVS Damage calculations

```{r}
tree_library = all_data_mvs
impact_dir = 'Python/MR_samples'
# sampling damage to each genus within each mvs
damage_samples_mvs <- list.files(path = impact_dir, pattern = "*.csv", full.names = T) |>
    map(function(x){
      # aggregate species impacts by genus
      impact_by_genus <- read.csv(x) |>
        group_by(Genus) |>
        summarise(Impact = mean(Damage, na.rm = T)) |>
        filter(!is.na(Impact)) |>
        rename(genus = Genus)
      # calculate impact to individual trees by mvs
      impact_mvs <- left_join(tree_library, impact_by_genus, by = c('genus')) |> # Assign based on genus match 
        mutate(Impact = if_else(is.na(Impact), 0, Impact, NA)) |> # if no info, impact = 0
        mutate(Impact_trees = trees_per_ha * Impact) |> # calc absolute impact to trees per ha
        group_by(mvsValue) |>
        # summarise(Impact_trees = sum(Impact_trees),
        #           Impact_prop = sum(Impact_trees) / sum(trees_per_ha)) |>
        st_drop_geometry() |>
        rename(MVS = mvsValue) |>
        ungroup() |>
        complete(MVS = c(1:75,79,80,90:93,96:99), fill = list(Impact_trees = 0, Impact = 0, trees_per_ha = 0))
      return(impact_mvs)
    })


# Damage_samples is the existing list of 1000 tibbles
# Average across 1000 trials

damage_avg_mvs <- damage_samples_mvs |>
  bind_rows() |>          # stack 1000
  group_by(MVS, genus, sizeclass) |>              # keep sizeclass
  summarise(
    Impact_mean       = mean(Impact, na.rm = TRUE),
    Impact_q025       = quantile(Impact, 0.025, na.rm = TRUE),
    Impact_q975       = quantile(Impact, 0.975, na.rm = TRUE),
    
    ImpactTrees_mean  = mean(Impact_trees, na.rm = TRUE),
    ImpactTrees_q025  = quantile(Impact_trees, 0.025, na.rm = TRUE),
    ImpactTrees_q975  = quantile(Impact_trees, 0.975, na.rm = TRUE),
    
    # keep stable attributes from the first run (identical across runs)
    trees_per_ha      = first(trees_per_ha),
    n_sites_per_mvs   = first(n_sites_per_mvs),
    .groups = "drop"
  ) |>
  arrange(MVS, genus, sizeclass)

# Output to a csv
write.csv(damage_avg_mvs, "Output/impact_avg_by_MVS_genus_sizeclass.csv", )
damage_avg_mvs <- read_csv("Output/impact_avg_by_MVS_genus_sizeclass.csv")

# Extract information on to a summarised MVS df
susceptibility_mvs <- damage_avg_mvs |>
  group_by(MVS) |>
  summarise(
    ImpactTrees_tot   = sum(ImpactTrees_mean, na.rm = TRUE),
    trees_per_ha_tot  = sum(trees_per_ha, na.rm = TRUE),
    Impact_proportion = (ImpactTrees_tot / trees_per_ha_tot),
    n_sites_per_mvs   = first(n_sites_per_mvs),
    .groups = "drop"
  ) |>
  replace_na(list(Impact_proportion = 0, n_sites_per_mvs = 0)) |> #data availability flag
  mutate(has_tree_data = if_else(n_sites_per_mvs > 0, TRUE, FALSE))

# Save MVGs with no tree data for overlay/masking
mvss_no_tree_data <- susceptibility_mvs |> 
  filter(has_tree_data == FALSE) |>
  dplyr::select(MVS)
```

#### Assigning descriptions to NVIS data and tabling

```{r}
# Set global default for NA string
set_flextable_defaults(na_str = "No data")

MVG_descs = c("Rainforests and Vine Thickets", #1
              "Eucalypt Tall Open Forests", #2
              "Eucalypt Open Forests", #3
              "Eucalypt Low Open Forests", #4
              "Eucalypt Woodlands", #5 
              "Acacia Forests and Woodlands", #6
              "Callitris Forests and Woodlands", #7
              "Casuarina Forests and Woodlands", #8
              "Melaleuca Forests and Woodlands", #9
              "Other Forests and Woodlands", #10
              "Eucalypt Open Woodlands", #11
              "Tropical Eucalypt Woodlands/Grasslands", #12
              "Acacia Open Woodlands", #13
              "Mallee Woodlands and Shrublands", #14
              "Low Closed Forests and Tall Closed Shrublands", #15
              "Acacia Shrublands", #16
              "Other Shrublands", #17
              "Heathlands", #18
              "Tussock Grasslands", #19
              "Hummock Grasslands", #20
              "Other Grasslands, Herblands, Sedgelands and Rushlands", #21
              "Chenopod Shrublands, Samphire Shrublands and Forblands", #22
              "Mangroves", #23
              "Inland Aquatic", #24
              "Cleared, non-native veg., buildings", #25
              "Unclassfied native veg.", #26
              "Naturally bare - sand, rock, claypan, mudflat", #27
              "Sea and estuaries", #28
              "Regrowth, modified native vegetation", #29
              "Unclassified forest", #30
              "Other Open Woodlands", #31
              "Mallee Open Woodlands and Sparse Mallee Shrublands", #32
              "No Data") #99)
names(MVG_descs) <- c(1:32,99)

# Format MVG summary
mvg_summary <- susceptibility_mvg |>
  mutate(
    ImpactTrees_tot = ifelse(has_tree_data == FALSE, NA, round(ImpactTrees_tot, digits = 1)),
    trees_per_ha_tot = ifelse(has_tree_data == FALSE, NA, round(trees_per_ha_tot, digits = 1)),
    Impact_proportion_pct = ifelse(has_tree_data == FALSE, NA, round(100*Impact_proportion, digits = 1)),
    Description = MVG_descs
  ) |>
    dplyr::select(-Impact_proportion)

# Check how many have no info
mvg_no_data <- susceptibility_mvg |>
  summarise(
    n_total = n(),
    n_false  = sum(has_tree_data == FALSE, na.rm = TRUE),
    percent_no_data = scales::percent(n_false / n_total, accuracy = 0.1)
  )

# Create flextable with description
ft_mvg <- mvg_summary |>
  dplyr::select(MVG, Description, ImpactTrees_tot, trees_per_ha_tot, Impact_proportion_pct) |>
  flextable() |>
  set_header_labels(
    MVG = "MVG",
    Description = "Description",
    ImpactTrees_tot = "Impacted plants/ha",
    trees_per_ha_tot = "Plants/ha",
    Impact_proportion_pct = "Impact proportion (%)"
  ) |>
  bold(part = "header") |>
  bg(part = "header", bg = "#D3D3D3") |>
  color(part = "header", color = "black") |>
  align(j = 3:5, align = "center") |>
  align(j = c("ImpactTrees_tot", "trees_per_ha_tot", "Impact_proportion_pct"),
        align = "center", part = "all") |>
  #font(fontname = "Georgia", part = "all") |>
  fontsize(size = 10, part = "all") |>
  theme_zebra(odd_body = "grey90") |>
  autofit() |>
  hline_top(part = "header", border = fp_border(color="black", width = 2)) |>
  hline_bottom(part = "header", border = fp_border(color="black", width = 2)) |>
  hline(i = 1:nrow(mvg_summary), border = fp_border(color="black", width = 1))

ft_mvg

# Export as high-res PNG
save_as_image(ft_mvg, path = "Output/images/mvg_impacts_table_with_descriptions.png", zoom = 2)

# Export to blank doc
doc <- read_docx()
doc <- body_add_flextable(doc, value = ft_mvg)
print(doc, target = "Output/plots/mvg_impacts.docx")

MVS_descs = c("Cool temperate rainforest",
              "Tropical or sub-tropical rainforest",
              "Eucalyptus (+/- tall) open forest with a dense broad-leaved and/or tree-fern understorey (wet sclerophyll)",
              "Eucalyptus open forests with a shrubby understorey",
              "Eucalyptus open forests with a grassy understorey",
              "Warm temperate rainforest",
              "Tropical Eucalyptus open forests and woodlands with a tall annual grassy understorey",
              "Eucalyptus woodlands with a shrubby understorey",
              "Eucalyptus woodlands with a tussock grass understorey",
              "Eucalyptus woodlands with a hummock grass understorey",
              "Tropical mixed spp forests and woodlands",
              "Callitris forests and woodlands",
              "Brigalow (Acacia harpophylla) forests and woodlands",
              "Other Acacia forests and woodlands",
              "Melaleuca open forests and woodlands",
              "Other forests and woodlands",
              "Boulders/rock with algae, lichen or scattered plants, or alpine fjaeldmarks",
              "Eucalyptus low open woodlands with hummock grass",
              "Eucalyptus low open woodlands with tussock grass",
              "Mulga (Acacia aneura) woodlands +/- tussock grass +/- forbs",
              "Other Acacia tall open shrublands and [tall] shrublands",
              "Acacia (+/- low) open woodlands and shrublands with chenopods",
              "Acacia (+/- low) open woodlands and shrublands with hummock grass",
              "Acacia (+/- low) open woodlands and shrublands +/- tussock grass",
              "Acacia (+/- low) open woodlands and sparse shrublands with a shrubby understorey",
              "Casuarina and Allocasuarina forests and woodlands",
              "Mallee with hummock grass",
              "Low closed forest or tall closed shrublands (including Acacia, Melaleuca and Banksia)",
              "Mallee with a dense shrubby understorey",
              "Heathlands",
              "Saltbush and bluebush shrublands",
              "Other shrublands",
              "Hummock grasslands",
              "Mitchell grass (Astrebla) tussock grasslands",
              "Blue grass (Dicanthium) and tall bunch grass (Vitiveria syn: Chrysopogon) tussock grasslands",
              "Temperate tussock grasslands",
              "Other tussock grasslands",
              "Wet tussock grassland with herbs, sedges or rushes, herblands or ferns",
              "Mixed chenopod, samphire +/- forbs",
              "Mangroves",
              "Saline or brackish sedgelands or grasslands",
              "Naturally bare, sand, rock, claypan, mudflat",
              "Salt lakes and lagoons",
              "Freshwater, dams, lakes, lagoons or aquatic plants",
              "Mulga (Acacia aneura) open woodlands and sparse shrublands +/- tussock grass",
              "Sea, estuaries (includes seagrass)",
              "Eucalyptus open woodlands with shrubby understorey",
              "Eucalyptus open woodlands with a grassy understorey",
              "Melaleuca shrublands and open shrublands",
              "Banksia woodlands",
              "Mulga (Acacia aneura) woodlands and shrublands with hummock grass",
              "Mulga (Acacia aneura) open woodlands and sparse shrublands with hummock grass",
              "Eucalyptus low open woodlands with a shrubby understorey",
              "Eucalyptus tall open forest with a fine-leaved shrubby understorey",
              "Mallee with an open shrubby understorey",
              "Eucalyptus (+/- low) open woodlands with a chenopod or samphire understorey",
              "Lignum shrublands and wetlands",
              "Leptospermum forests and woodlands",
              "Eucalyptus woodlands with ferns, herbs, sedges, rushes or wet tussock grassland",
              "Eucalyptus tall open forests and open forests with ferns, herbs, sedges, rushes or wet tussock grasses",
              "Mallee with a tussock grass understorey",
              "Dry rainforest or vine thickets",
              "Sedgelands, rushes or reeds",
              "Other grasslands",
              "Eucalyptus woodlands with a chenopod or samphire understorey",
              "Open mallee woodlands and sparse mallee shrublands with a hummock grass understorey",
              "Open mallee woodlands and sparse mallee shrublands with a tussock grass understorey",
              "Open mallee woodlands and sparse mallee shrublands with an open shrubby understorey",
              "Open mallee woodlands and sparse mallee shrublands with a dense shrubby understorey",
              "Callitris open woodlands",
              "Casuarina and Allocasuarina open woodlands with a tussock grass understorey",
              "Casuarina and Allocasuarina open woodlands with a hummock grass understorey",
              "Casuarina and Allocasuarina open woodlands with a chenopod shrub understorey",
              "Casuarina and Allocasuarina open woodlands with a shrubby understorey",
              "Melaleuca open woodlands",
              "Other open woodlands",
              "Other sparse shrublands and sparse heathlands",
              "Regrowth or modified forests and woodlands",
              "Regrowth or modified shrublands",
              "Regrowth or modified graminoids",
              "Regrowth or modified chenopod shrublands, samphire or forblands",
              "Unclassified forest",
              "Unclassified native vegetation",
              "Cleared, non-native vegetation, buildings",
              "Unknown/no data"
              )
names(MVS_descs) <- c(1:75,79,80,90:93,96:99)

# Format MVS summary
mvs_summary <- susceptibility_mvs |>
  mutate(
    ImpactTrees_tot = ifelse(has_tree_data == FALSE, NA, round(ImpactTrees_tot, digits = 1)),
    trees_per_ha_tot = ifelse(has_tree_data == FALSE, NA, round(trees_per_ha_tot, digits = 1)),
    Impact_proportion_pct = ifelse(has_tree_data == FALSE, NA, round(100*Impact_proportion, digits = 1)),
    Description = MVS_descs
  ) |>
    dplyr::select(-Impact_proportion)

# Check how many have no info
mvs_no_data <- susceptibility_mvs |>
  summarise(
    n_total = n(),
    n_false  = sum(has_tree_data == FALSE, na.rm = TRUE),
    percent_no_data = scales::percent(n_false / n_total, accuracy = 0.1)
  )


# Create flextable with description
ft_mvs <- mvs_summary |>
  dplyr::select(MVS, Description, ImpactTrees_tot, trees_per_ha_tot, Impact_proportion_pct) |>
  flextable() |>
  set_header_labels(
    MVS = "MVS",
    Description = "Description",
    ImpactTrees_tot = "Impacted plants/ha",
    trees_per_ha_tot = "Plants/ha",
    Impact_proportion_pct = "Impact proportion (%)"
  ) |>
  bold(part = "header") |>
  bg(part = "header", bg = "#D3D3D3") |>
  color(part = "header", color = "black") |>
  align(j = 3:5, align = "center") |>
  align(j = c("ImpactTrees_tot", "trees_per_ha_tot", "Impact_proportion_pct"),
        align = "center", part = "all") |>
  fontsize(size = 10, part = "all") |>
  theme_zebra(odd_body = "grey90") |>
  autofit() |>
  hline_top(part = "header", border = fp_border(color="black", width = 2)) |>
  hline_bottom(part = "header", border = fp_border(color="black", width = 2)) |>
  hline(i = 1:nrow(mvs_summary), border = fp_border(color="black", width = 1))

ft_mvs

# Export as high-res PNG
save_as_image(ft_mvs, path = "Output/images/mvs_impacts_table_with_descriptions.png", zoom = 2)

# Export to blank doc
doc <- read_docx()
doc <- body_add_flextable(doc, value = ft_mvs)
print(doc, target = "Output/plots/mvs_impacts.docx")
```

#### Assign data to rasters

First, rasters are loaded in and MVGs and MVSs are assigned impact proportions taken from the susceptibility tables

```{r}
# Setting up rasters in terra
MVG_raster <- rast("data/spatial/NVIS_MVG_RAS_AA.tif")
MVS_raster <- rast("data/spatial/NVIS_MVS_RAS_AA.tif")

# Assign levels based on NVIS descriptions
levels(MVG_raster) = data.frame(MVG=c(1:32,99), description=MVG_descs)

levels(MVS_raster) = data.frame(MVS=c(1:75,79,80,90:93,96:99), description=MVS_descs)

# Aggregate susceptibility map (MVG-level)
# Lookup table for impact proportion
MVG_lut <- data.frame(from = susceptibility_mvg$MVG,
                      to   = susceptibility_mvg$Impact_proportion)


# Create a quick n dirty susceptibility raster by MVG
MVG_Impact_raster <- classify(MVG_raster, rcl = MVG_lut)


# Aggregate susceptibility map (MVS-level)
# Lookup table for impact proportion
MVS_lut <- data.frame(from = susceptibility_mvs$MVS,
                      to   = susceptibility_mvs$Impact_proportion)


# Create a quick n dirty susceptibility raster by MVS
MVS_Impact_raster <- classify(MVS_raster, rcl = MVS_lut)

# Plot these side by side
# Stack rasters for visualisation
MVG_MVS_stack <- c(MVG_Impact_raster, MVS_Impact_raster)
names(MVG_MVS_stack) <- c("Susceptiblity by MVG","Susceptibility by MVS")


```

## Combining with suitability

Once the suitability is modeled using MaxEnt and the susceptibility is calculated and projected spatially using NVIS information, these two rasters are mutliplied to give a relative probability of A. psidii presence in a given cell. This is possible as the suitability raster output from maxent produces a relative probability field with values 0 through to 1 which can be interpreted as a probability that a given cell is suitable. The susceptibility rasters on the other hand provide an estimation of the proportion of plants (or plant material perhaps) that is assumed to be damaged by myrtle rust in a given cell. As this is also a range from 0 to 1, combining these two rasters will provide an index from 0 through to 1 which suggests the expected level of damage if myrtle rust were to occur at a given site, based on the available data.

```{r}

#Load in suitability raster
suit <- rast("data/Predictions/maxent_prediction.tif")

# Apply threshold for categorisation later - using cloglog threshold from MTSS
suit_thresh <- suit
suit_thresh[suit_thresh <= 0.114] <- 0


# Make sure CRS matches
MVG_MVS_stack <- project(MVG_MVS_stack, suit)

# Resampling to ensur ext and res match
MVG_MVS_stack_align <- resample(MVG_MVS_stack, suit, method = "bilinear") # both rasters are continous data so bilinear (could potentially switch to cubic)

# Multiply susceptibility stack by suitability
sus_x_suit <- MVG_MVS_stack_align * suit 
names(sus_x_suit) <- c("Mrust_val_MVG", "Mrust_val_MVS")

writeRaster(sus_x_suit, filename = "Output/data/Apsidii_sus_x_suit.tif", overwrite = TRUE)
sus_x_suit <- rast("Output/data/Apsidii_sus_x_suit.tif")
names(sus_x_suit) <- c("Mrust_val_MVG", "Mrust_val_MVS") #in case names not carryin in

plot(sus_x_suit)

# Now multiply again but with thresholded 
sxs_thresh <- MVG_MVS_stack_align * suit_thresh
names(sxs_thresh) <- c("Mrust_val_MVG", "Mrust_val_MVS")

writeRaster(sxs_thresh, filename = "Output/data/Apsidii_sxs_thresh.tif")
sxs_thresh <- rast("Output/data/Apsidii_sxs_thresh.tif")

# Scaling the mvs raster now to get a value between 0 and 100
mm_mvs <- global(sus_x_suit$Mrust_val_MVS, c("min","max"), na.rm = TRUE, wide = TRUE)
mins_mvs <- mm_mvs$min   # true min per mvs layer
maxs_mvs <- mm_mvs$max   # true max per mvs layer

sus_x_suit_mvs_scaled <- stretch(sus_x_suit$Mrust_val_MVS, 
                             minv=0, 
                             maxv=100, 
                             smin=mins_mvs, 
                             smax=maxs_mvs)
writeRaster(sus_x_suit_mvs_scaled, "Output/data/Apsidii_sus_x_suit_mvs_scaled_0100.tif")


#scaling the mvg raster
mm_mvg <- global(sus_x_suit$Mrust_val_MVG, c("min","max"), na.rm = TRUE, wide = TRUE)
mins_mvg <- mm_mvg$min   # true min per mvg layer
maxs_mvg <- mm_mvg$max   # true max per mvg layer

sus_x_suit_mvg_scaled <- stretch(sus_x_suit$Mrust_val_MVG, 
                             minv=0, 
                             maxv=100, 
                             smin=mins_mvg, 
                             smax=maxs_mvg)
writeRaster(sus_x_suit_mvg_scaled, "Output/data/Apsidii_sus_x_suit_mvg_scaled_0100.tif", overwrite = TRUE)

# scaling the thresholded raster
mm_mvs <- global(sxs_thresh$Mrust_val_MVS, c("min","max"), na.rm = TRUE, wide = TRUE)
mins_mvs <- mm_mvs$min   # true min per mvs layer
maxs_mvs <- mm_mvs$max   # true max per mvs layer

sxs_thresh_scaled <- stretch(sxs_thresh$Mrust_val_MVS, 
                             minv=0, 
                             maxv=100, 
                             smin=mins_mvs, 
                             smax=maxs_mvs)

# scaling the thresholded raster
mm_mvg <- global(sxs_thresh$Mrust_val_MVG, c("min","max"), na.rm = TRUE, wide = TRUE)
mins_mvg <- mm_mvg$min   # true min per mvg layer
maxs_mvg <- mm_mvg$max   # true max per mvg layer

sxs_thresh_scaled_mvg <- stretch(sxs_thresh$Mrust_val_MVG, 
                             minv=0, 
                             maxv=100, 
                             smin=mins_mvg, 
                             smax=maxs_mvg)


writeRaster(sxs_thresh_scaled, "Output/data/Apsidii_sxs_thresh_scaled_0100.tif", overwrite = TRUE)
writeRaster(sxs_thresh_scaled_mvg, "Output/data/Apsidii_sxs_thresh_mvg_scaled_0100.tif", overwrite = TRUE)
```

### Mask layer of missing values

```{r}
# ---- Create mask for MVGs with no tree data ----
# downsample
# Make sure CRS matches
MVG_raster <- project(MVG_raster, suit)

# Resampling to ensur ext and res match
MVG_align <- resample(MVG_raster, suit, method = "bilinear")

# MVG raster is categorical
lvl <- levels(MVG_raster)[[1]]   # ID = MVG column, description

# Map categories to numeric values (ID)
MVG_raster_numeric <- classify(MVG_align, rcl = cbind(lvl$MVG, lvl$MVG), others = NA)

# mvgs_no_tree_data is a vector of MVG values missing tree data
missing_mvgs <- mvgs_no_tree_data$MVG

# mask raster: 1 where MVG is missing, NA otherwise
mvg_no_data_mask <- ifel(MVG_raster_numeric %in% missing_mvgs, 1, NA)
# Project to an equal areas crs
mvg_no_data_alb <- project(mvg_no_data_mask, "EPSG:3577")

# Build a small data.frame with xy and value
mvg_no_data_mask_df <- as.data.frame(mvg_no_data_mask, xy = TRUE, na.rm = TRUE)

write_csv(mvg_no_data_mask_df, "data/processed/mvg_no_data_mask.csv")

# ---- Create mask for MVSs with no tree data ----
# downsample
# Make sure CRS matches
MVS_raster <- project(MVS_raster, suit)

# Resampling to ensure ext and res match
MVS_align <- resample(MVS_raster, suit, method = "bilinear")

# MVS raster is categorical
lvl <- levels(MVS_raster)[[1]]   # ID = MVG column, description

# Map categories to numeric values (ID)
MVS_raster_numeric <- classify(MVS_align, rcl = cbind(lvl$MVS, lvl$MVS), others = NA)

# mvss_no_tree_data is a vector of MVS values missing tree data
missing_mvss <- mvss_no_tree_data$MVS

# mask raster: 1 where MVG is missing, NA otherwise
mvs_no_data_mask <- ifel(MVS_raster_numeric %in% missing_mvss, 1, NA)
# # Project to an equal areas crs - got rid of this
# mvs_no_data_alb <- project(mvs_no_data_mask, "EPSG:3577")

# Build a small data.frame with xy and value
mvs_no_data_mask_df <- as.data.frame(mvs_no_data_mask, xy = TRUE, na.rm = TRUE)

write_csv(mvs_no_data_mask_df, "data/processed/mvs_no_data_mask.csv")
```

### Visualisations

```{r}
# Australia outline for maps
aus_alb_shp <- st_transform(ne_countries(country = "australia", scale = "medium", returnclass = "sf"), "EPSG:3577")
aus_main_shp <- st_crop(aus_alb_shp,       
                    xmin = -1.888e6, xmax = 5.5e6,
                    ymin = -5e6, ymax = -1e6)
aus_main_shp_wgs84 <- st_transform(aus_main_shp, crs = "EPSG:4326")

# --- susceptibility rasters with masks
# Ensure they're in the same CRS first
MVG_Impact_aligned <- project(MVG_Impact_raster, mvg_no_data_mask)
MVS_Impact_aligned <- project(MVS_Impact_raster, mvs_no_data_mask)

# Resample to match resolutions 
MVG_Impact_resampled <- resample(MVG_Impact_aligned, mvg_no_data_mask, method = "bilinear")
MVS_Impact_resampled <- resample(MVS_Impact_aligned, mvs_no_data_mask, method = "bilinear")

# Clip extent to the same as the shapefile
MVG_Impact_crop <- crop(MVG_Impact_resampled, aus_main_shp_wgs84)
MVS_Impact_crop <- crop(MVS_Impact_resampled, aus_main_shp_wgs84)


# Add an id column so ggplot treats each cell separately
mvs_no_data_mask_df$mask <- "No data"
mvs_no_data_mask_df$mask <- "No data"

# Convert the two impact rasters to data frames (with x, y, and value)
MVG_df <- as.data.frame(MVG_Impact_crop, xy = TRUE, na.rm = FALSE)
names(MVG_df)[3] <- "Impact"

MVS_df <- as.data.frame(MVS_Impact_crop, xy = TRUE, na.rm = FALSE)
names(MVS_df)[3] <- "Impact"

# Find common color scale range
common_limits <- range(c(MVG_df$Impact, MVS_df$Impact), na.rm = TRUE)

# MVG map
p_mvg <- ggplot() +
  geom_raster(data = MVG_df, aes(x, y, fill = Impact)) +
  geom_raster(data = mvg_no_data_mask_df, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  geom_sf(data = aus_main_shp_wgs84, fill = NA, colour = "black", size = 0.3) +
  scale_fill_viridis_c(
    option = "viridis",
    name = "Impact",
    na.value = "transparent"
  ) +
  # Alpha legend for no data
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  coord_sf(expand = TRUE) +
  labs(title = "Estimated impact by MVG") +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

# MVS map
p_mvs <- ggplot() +
  geom_raster(data = MVS_df, aes(x, y, fill = Impact)) +
  geom_raster(data = mvs_no_data_mask_df, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  geom_sf(data = aus_main_shp_wgs84, fill = NA, colour = "black", size = 0.3) +
  scale_fill_viridis_c(
    option = "viridis",
    name = "Impact",
    na.value = "transparent"
  ) +
  # Alpha legend for no data
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  coord_sf(expand = TRUE) +
  labs(title = "Estimated impact by MVS") +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))


combined_plot <- p_mvg + p_mvs + plot_layout(guides = "collect") &
  theme(legend.position = "bottom")


combined_plot
ggsave("Output/plots/MVG_MVS_susceptibility_comparison.tiff",
       combined_plot, width = 12.5, height = 6, dpi = 300)

# --- susceptibility suitability combined

# load in scaled rast
sus_x_suit_mvs_scaled <- rast("Output/data/Apsidii_sus_x_suit_mvs_scaled_0100.tif")
sus_x_suit_mvg_scaled <- rast("Output/data/Apsidii_sus_x_suit_mvg_scaled_0100.tif")


# CUTOFF RASTER SUPERCEDED BY THRESHOLDED RASTER
# # Cut off raster values below 1 to highlight only important parts
# cutoff_rast <- ifel(sus_x_suit_mvs_scaled<1, NA, sus_x_suit_mvs_scaled)
# 
# writeRaster(cutoff_rast, "Output/data/Apsidii_sus_x_suit_mvs_scaled_0100_cutoff.tif")
# cutoff_rast <- rast("Output/data/Apsidii_sus_x_suit_mvs_scaled_0100_cutoff.tif")

# Project to an equal areas crs
sxs_alb <- project(sus_x_suit_mvs_scaled, "EPSG:3577")
writeRaster(sxs_alb, "data/processed/sxs_alb.tif")
sxs_alb <- rast("data/processed/sxs_alb.tif")
sxs_mvg_alb <- project(sus_x_suit_mvg_scaled, "EPSG:3577")

# make sus x suit data frames
df_sxs <- as.data.frame(sxs_alb, xy = TRUE, na.rm = TRUE)
df_sxs$x <- round(df_sxs$x, 0)  # EPSG:3577 is in metres, so round to whole metres
df_sxs$y <- round(df_sxs$y, 0)
write_csv(df_sxs, "data/processed/df_sxs_alb.csv")

df_sxs_mvg <- as.data.frame(sxs_mvg_alb, xy = TRUE, na.rm = TRUE)
df_sxs_mvg$x <- round(df_sxs_mvg$x, 0)  # EPSG:3577 is in metres, so round to whole metres
df_sxs_mvg$y <- round(df_sxs_mvg$y, 0)
write_csv(df_sxs_mvg, "data/processed/df_sxs_mvg_alb.csv")




# Plot scaled myrtle rust "danger" across australia
mvs_sxs_map <- ggplot() +
  geom_raster(data = df_sxs, aes(x, y, fill = Mrust_val_MVS)) +
  geom_sf(data = aus_main_shp, fill = NA, colour = "black", size = 0.3) +
  geom_raster(data = mvs_no_data_mask_df, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  scale_fill_viridis_c(
    option = "inferno",
    direction = -1,
    values = scales::rescale(c(0, 5, 20, 50, 70, 100)),
    name = "Relative myrtle \nrust intensity",
    na.value = "transparent"
  ) +
  # Alpha legend for no data
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  coord_sf(expand = TRUE) +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank())

mvs_sxs_map
ggsave("Output/images/mvg_sxs_map.png", plot = sxs_map, width = 10, height = 8, dpi = 600)

# Add a variable for "mask" fill
mvg_no_data_mask_df$mask <- "No data"

# Plot scaled myrtle rust "danger" across australia
mvg_sxs_map <- ggplot() +
  geom_raster(data = df_sxs_mvg, aes(x, y, fill = Mrust_val_MVG)) +
  geom_sf(data = aus_main_shp, fill = NA, colour = "black", size = 0.3) +
  geom_raster(data = mvg_no_data_mask_df, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  scale_fill_viridis_c(
    option = "inferno",
    direction = -1,
    values = scales::rescale(c(0, 5, 20, 50, 70, 100)),
    name = "Relative myrtle \nrust intensity",
    na.value = "transparent",
    guide = guide_colorbar(order = 1)
  ) +
  # Alpha legend for no data
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  coord_sf() +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank())

mvg_sxs_map
ggsave("Output/images/mvg_sxs_map.png", plot = sxs_map, width = 8, height = 6, dpi = 600)


```

## Categorisation of Myrtle rust raster

Doing this just for visualisation purposes and for making generalisations later.

```{r}
# Load in thresholded raster
sxs_thresh_scaled <- rast("Output/data/Apsidii_sxs_thresh_scaled_0100.tif")

# Define breaks for non-zero values using Jenks/Fisher method
nonzero_vals <- values(sxs_thresh_scaled)[!is.na(values(sxs_thresh_scaled)) > 0]
fisher_breaks <- classIntervals(nonzero_vals, n = 4, style = "fisher", samp_prop = 0.5)$brks


# Classify the susceptiblity times suitability raster
sus_x_suit_class <- app(sxs_thresh_scaled, fun = function(x) {
  ifelse(x == 0, 0, #no risk is below 1
         findInterval(x, fisher_breaks, rightmost.closed = TRUE))
})
writeRaster(sus_x_suit_class, "Output/data/sus_x_suit_class.tif", overwrite = TRUE)
sus_x_suit_class <- rast("Output/data/sus_x_suit_class.tif")

names(sus_x_suit_class) <- "Hazard Category"
levels(sus_x_suit_class) <- data.frame(value = 0:4, riskcat = c("Very Low", "Low", "Moderate", "High", "Very High"))



```

### Sensitivity analysis

```{r}
# Classification methods to test
methods <- c("fisher", "jenks", "equal")

# Function to classify raster
classify_raster_iteration <- function(method, iter) {
  
  # Number of non-zero classes
  n_classes <- 4
  
  # Compute breaks
  if (method %in% c("fisher","jenks")) {
    breaks <- classIntervals(nonzero_vals, n = n_classes, style = method, samp_prop = 0.5)$brks
  } else if (method == "equal") {
    breaks <- classIntervals(nonzero_vals, n = n_classes, style = "equal")$brks
  }
  
  # Classify raster
  classified <- app(sxs_thresh_scaled, function(x) {
    x[is.na(x) | x == 0] <- 0
    x[x > 0] <- findInterval(x[x > 0], breaks, rightmost.closed = TRUE)
    return(x)
  })
  
  # Count cells
  counts <- as.data.frame(terra::freq(classified))
  counts$iteration <- iter
  counts$method <- method
  return(counts)
}

# Methods and iterations
methods <- c("fisher", "jenks", "equal")
n_iter <- 20

# Run iterations for all methods
sensitivity_df <- map_dfr(methods, function(m) {
  map_dfr(1:n_iter, ~ classify_raster_iteration(m, .x))
})

# Label risk categories
sensitivity_summary <- sensitivity_df |>
  mutate(
    riskcat = factor(value, levels = 0:4,
                     labels = c("Very Low","Low","Moderate","High","Very High"))
  ) |>
  group_by(method, iteration, riskcat) |>
  summarise(cell_count = sum(count), .groups = "drop") |>
  group_by(method, iteration) |>
  mutate(percentage = 100 * cell_count / sum(cell_count))

# Compute variance across iterations per method
class_variance <- sensitivity_summary |>
  group_by(method, riskcat) |>
  summarise(
    mean_pct = mean(percentage),
    sd_pct = sd(percentage),
    cv_pct = sd_pct / mean_pct * 100
  )

# Mean % per category per method
method_means <- sensitivity_summary |>
  group_by(method, riskcat) |>
  summarise(mean_pct = mean(percentage), .groups = "drop")

method_variance <- method_means |>
  group_by(riskcat) |>
  summarise(
    mean_across_methods = mean(mean_pct),
    sd_across_methods = sd(mean_pct),
    cv_across_methods = sd_across_methods / mean_across_methods * 100
  )

# Filter out verylow
means_noverylow <- method_means |>
  filter(riskcat != 'Very Low')
var_noverylow <- class_variance |>
  filter(riskcat != 'Very Low')
sensitivity_summary_filtered <- sensitivity_summary |>
  filter(riskcat != 'Very Low')

# Plot
ggplot(means_noverylow, aes(x = riskcat, y = mean_pct, fill = method)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(
    data = var_noverylow,
    aes(x = riskcat, ymin = mean_pct - sd_pct, ymax = mean_pct + sd_pct),
    width = 0.2,
    color = "black",
    position = position_dodge(width = 0.8)  # match bar dodge
  ) +
  labs(
    x = "Hazard Category",
    y = "Mean % of Cells",
    title = "Variance Between Hazard Classification Methods (Excluding Very Low)",
    fill = "Method"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("Output/plots/class_method_sensitivity_col.tiff", width = 10, height = 8, dpi = 300)
# Boxplot
ggplot(sensitivity_summary_filtered |> filter(method != "equal"), aes(x = riskcat, y = percentage, fill = method)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +
  labs(
    x = "Hazard Category",
    y = "% of Cells",
    title = "Distribution of Cell Percentages Across Methods and Iterations",
    fill = "Method"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("Output/plots/class_method_sensitivity_box.tiff", width = 10, height = 8, dpi = 300)

```

### Clipping to states

```{r}
# ABS States and Territories - 2021 - for state boundaries https://www.abs.gov.au/statistics/standards/australian-statistical-geography-standard-asgs-edition-3/jul2021-jun2026/access-and-downloads/digital-boundary-files?utm_source=chatgpt.com
aus_states <- st_read("data/spatial/STE_2021_AUST_GDA2020.shp")
aus_states <- st_transform(aus_states, crs(sus_x_suit_class))

#State selections
vic_shp <- aus_states |> 
  filter(STE_CODE21 == 2) |>
  summarise(geometry = st_union(geometry))

nsw_shp <- aus_states |> 
  filter(STE_CODE21 %in% c(1,8)) |>
  summarise(geometry = st_union(geometry))
# Split multipolygon into individual polygons
nsw_parts <- st_cast(nsw_shp, "POLYGON")
# Remove lord howe island
nsw_parts <- nsw_parts |>
  mutate(
    centroid = st_centroid(geometry),
    lon = st_coordinates(centroid)[,1],
    lat = st_coordinates(centroid)[,2]
  )
nsw_mainland <- nsw_parts |>
  filter(lon < 155) |>  # can adjust based on CRS
  summarise(geometry = st_union(geometry))
  
qld_shp <- aus_states |> 
  filter(STE_CODE21 == 3) |>
  summarise(geometry = st_union(geometry))

tas_shp <- aus_states |> 
  filter(STE_CODE21 == 6)|>
  summarise(geometry = st_union(geometry))

nt_shp <- aus_states |> filter(STE_CODE21 == 7) |>
  summarise(geometry = st_union(geometry))

# Do eastern states
# Add identifiers
nsw_mainland$state <- "NSW and ACT"
vic_shp$state <- "VIC"
qld_shp$state <- "QLD"
tas_shp$state <- "TAS"
nt_shp$state <- "NT"

qldnsw_shp <- bind_rows(nsw_mainland, qld_shp)
east_states_shp <- bind_rows(nsw_mainland, vic_shp, qld_shp, tas_shp)
east_nt_states_shp <- bind_rows(nsw_mainland, vic_shp, qld_shp, tas_shp, nt_shp)

st_write(east_states_shp, "data/processed/east_states_shp.shp")

# No data raster
mvs_no_data_mask_df <- read_csv("data/processed/mvs_no_data_mask.csv")

# Crop and mask using terra
# Vic
vic_sxs <- crop(sus_x_suit_class, vic_shp)
vic_sxs_clip <- mask(vic_sxs, vic_shp)


vic_no_data <- crop(mvs_no_data_mask, vic_shp)
vic_no_data_clip <- mask(vic_no_data, vic_shp)

# NSW
nsw_sxs <- crop(sus_x_suit_class, nsw_mainland)
nsw_sxs_clip <- mask(nsw_sxs, nsw_mainland)

nsw_no_data <- crop(mvs_no_data_mask, nsw_mainland)
nsw_no_data_clip <- mask(nsw_no_data, nsw_mainland)

# QLD 
qld_sxs <- crop(sus_x_suit_class, qld_shp)
qld_sxs_clip <- mask(qld_sxs, qld_shp)

qld_no_data <- crop(mvs_no_data_mask, qld_shp)
qld_no_data_clip <- mask(qld_no_data, qld_shp)

# Tas
tas_sxs <- crop(sus_x_suit_class, tas_shp)
tas_sxs_clip <- mask(tas_sxs, tas_shp)

tas_no_data <- crop(mvs_no_data_mask, tas_shp)
tas_no_data_clip <- mask(tas_no_data, tas_shp)

# NT 
nt_sxs <- crop(sus_x_suit_class, nt_shp)
nt_sxs_clip <- mask(nt_sxs, nt_shp)

nt_no_data <- crop(mvs_no_data_mask, nt_shp)
nt_no_data_clip <- mask(nt_no_data, nt_shp)

# East
east_sxs <- crop(sus_x_suit_class, east_states_shp)
east_sxs_clip <- mask(east_sxs, east_states_shp)

east_no_data <- crop(mvs_no_data_mask, east_states_shp)
east_no_data_clip <- mask(east_no_data, east_states_shp)
```

### Categorical visualisations

```{r}

## --- Some visualisations ---


# make class df
df_sxs_class <- as.data.frame(sus_x_suit_class, xy = TRUE, na.rm = TRUE)

write_csv(df_sxs_class, "data/processed/df_sxs_class.csv")

# Mask
mvs_no_data_mask_df <- read_csv("data/processed/mvs_no_data_mask.csv")

# Aus 
aus_main_shp <- st_transform(aus_main_shp, crs = crs(sus_x_suit_class))

# --- Australia map ---
# Plot scaled myrtle rust "danger" across australia
sxs_class_map <- ggplot() +
  geom_raster(data = df_sxs_class, aes(x, y, fill = riskcat)) +
  geom_raster(data = mvs_no_data_mask_df, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  geom_sf(data = aus_main_shp, fill = NA, colour = "black", linewidth = 0.3) +
  scale_fill_manual(
    values = c("Very Low" = "#56B4E9",   # sky blue
             "Low"     = "#009E73",   # bluish green
             "Moderate"= "#F0E442",   # yellow
             "High"    = "#E69F00",   # orange
             "Very High" = "#D55E00"  # vermillion
               ),
    name = "Hazard category",
    na.value = "transparent"
  ) +
  
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  
  
  coord_sf() +
  #annotation_scale(location = "bl", width_hint = 0.3) +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank())

sxs_class_map

# --- vic map ---

# dataframe
df_class_vic <- as.data.frame(vic_sxs_clip, xy = TRUE, na.rm = TRUE)
df_no_data_vic <- as.data.frame(vic_no_data_clip, xy = TRUE, na.rm = TRUE)


vic_class_map <- ggplot() +
  geom_raster(data = df_class_vic, aes(x, y, fill = riskcat)) +
  geom_raster(data = df_no_data_vic, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  geom_sf(data = vic_shp, fill = NA, colour = "black", linewidth = 0.6) +
  scale_fill_manual(
    values = c("Very Low" = "#56B4E9",   # sky blue
             "Low"     = "#009E73",   # bluish green
             "Moderate"= "#F0E442",   # yellow
             "High"    = "#E69F00",   # orange
             "Very High" = "#D55E00"  # vermillion
               ),
    name = "Hazard category",
    na.value = "transparent"
  ) +
  
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  
  
  coord_sf() +
  annotation_north_arrow(location = "tr", height = unit(1, "in"), width = unit(1, "in"),
                         which_north = "true", style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

vic_class_map

# --- tas map ---

# dataframe
df_class_tas <- as.data.frame(tas_sxs_clip, xy = TRUE, na.rm = TRUE)
df_no_data_tas <- as.data.frame(tas_no_data_clip, xy = TRUE, na.rm = TRUE)

tas_class_map <- ggplot() +
  geom_raster(data = df_class_tas, aes(x, y, fill = riskcat)) +
  geom_raster(data = df_no_data_tas, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  geom_sf(data = tas_shp, fill = NA, colour = "black", linewidth = 0.6) +
  scale_fill_manual(
    values = c("Very Low" = "#56B4E9",   # sky blue
             "Low"     = "#009E73",   # bluish green
             "Moderate"= "#F0E442",   # yellow
             "High"    = "#E69F00",   # orange
             "Very High" = "#D55E00"  # vermillion
               ),
    name = "Hazard category",
    na.value = "transparent"
  ) +
  
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  
  
  coord_sf() +
  annotation_north_arrow(location = "tr", height = unit(1, "in"), width = unit(1, "in"),
                         which_north = "true", style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

tas_class_map

# --- qld map ---

# dataframe
df_class_qld <- as.data.frame(qld_sxs_clip, xy = TRUE, na.rm = TRUE)
df_no_data_qld <- as.data.frame(qld_no_data_clip, xy = TRUE, na.rm = TRUE)



qld_class_map <- ggplot() +
  geom_raster(data = df_class_qld, aes(x, y, fill = riskcat)) +
  geom_raster(data = df_no_data_qld, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  geom_sf(data = qld_shp, fill = NA, colour = "black", linewidth = 0.3) +
  scale_fill_manual(
    values = c("Very Low" = "#56B4E9",   # sky blue
             "Low"     = "#009E73",   # bluish green
             "Moderate"= "#F0E442",   # yellow
             "High"    = "#E69F00",   # orange
             "Very High" = "#D55E00"  # vermillion
               ),
    name = "Hazard category",
    na.value = "transparent"
  ) +
  
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  
  
  coord_sf() +
  annotation_north_arrow(location = "tr", height = unit(1, "in"), width = unit(1, "in"), which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

qld_class_map


# --- nsw map ---

# dataframe
df_class_nsw <- as.data.frame(nsw_sxs_clip, xy = TRUE, na.rm = TRUE)
df_no_data_nsw <- as.data.frame(nsw_no_data_clip, xy = TRUE, na.rm = TRUE)



nsw_class_map <- ggplot() +
  geom_raster(data = df_class_nsw, aes(x, y, fill = riskcat)) +
  geom_raster(data = df_no_data_nsw, aes(x = x, y = y, alpha = "No data"), fill = "grey80") +
  geom_sf(data = nsw_mainland, fill = NA, colour = "black", linewidth = 0.3) +
  scale_fill_manual(
    values = c("Very Low" = "#56B4E9",   # sky blue
             "Low"     = "#009E73",   # bluish green
             "Moderate"= "#F0E442",   # yellow
             "High"    = "#E69F00",   # orange
             "Very High" = "#D55E00"  # vermillion
               ),
    name = "Hazard category",
    na.value = "transparent"
  ) +
  
  scale_alpha_manual(
    values = c("No data" = 0.8),
    name = NULL
  ) +
  
  
  coord_sf() +
  #annotation_scale(location = "bl", width_hint = 0.3, style = "ticks", unit_category = "metric") +
  annotation_north_arrow(location = "tl", height = unit(1, "in"), width = unit(1, "in"),
                         which_north = "true", style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

nsw_class_map
```

# Risk assessment

This portion of the code is the risk assessment. Mammal exposure and vulnerability to myrtle rust is calculated here.

## Proportion of endangered mammal habitat in myrtle rust

```{r}
# Read fauna geodatabase file
fauna_sf <- st_read("data/raw/snes_public_gdb.gdb")

# TERRA version
gdb <- vect("data/raw/snes_public_gdb.gdb")
gdb_transform <- project(gdb, crs(sus_x_suit_class))

# Put in same CRS as everything else
fauna_sf <- st_transform(fauna_sf, crs(sus_x_suit_mvs_scaled))

# Filter to only have terrestrial mammals
mammals_sf <- fauna_sf |> 
  filter(TAXON_CLASS == "Mammalia") |>
  filter(!TAXON_ORDER %in% c("Cetacea","Sirenia","Pinnipedia")) 

# Adding a row index to help with calculations
mammals_sf$ID <- seq_len(nrow(mammals_sf))

# Extract raster values per polygon within gdb
vals_list <- terra::extract(sus_x_suit_mvs_scaled, vect(mammals_sf))

index_list <- terra::extract(sus_x_suit_class, vect(mammals_sf))

write_csv(vals_list, "data/processed/mamm_val_df.csv", col_names = TRUE)
vals_list <- read_csv("data/processed/mamm_val_df.csv")

write_csv(index_list, "data/processed/mamm_index_df.csv", col_names = TRUE)

# Calculate counts for each cell value in polygons
# DEPRECATED
# counts_MVG <- vals_list |>
  # group_by(ID, Mrust_val_MVG) |>
  # summarise(count = n()) |>
  # ungroup()

counts_MVS <- vals_list |>
  group_by(ID, Mrust_val_MVS) |>
  summarise(count = n()) |>
  ungroup()

# Do it for the categories
indices_MVS <- index_list |>
  group_by(ID, riskcat) |>
  summarise(count = n()) |>
  ungroup()

# Calculate proportion of habitat with non-zero myrtle rust impact + mean value of myrtle rust exposure within the range
# DEPRECATED
# props_MVG <- counts_MVG |>
#   group_by(ID) |>
#   summarise(
#     total = sum(count, na.rm = TRUE),
#     nonzero = sum(count[`Myrtle rust index MVG` != 0], na.rm = TRUE),
#     prop_nonzero = nonzero / total,
#     weighted_mean = weighted.mean(`Myrtle rust index MVG`, count, na.rm = TRUE)
#   )

props_MVS <- counts_MVS |>
  group_by(ID) |>
  summarise(
    total = sum(count, na.rm = TRUE),
    nonzero = sum(count[Mrust_val_MVS != 0], na.rm = TRUE),
    prop_nonzero = nonzero / total,
    weighted_mean = weighted.mean(Mrust_val_MVS, count, na.rm = TRUE)
  )
#####
# For the index (ONLY FOR MVS)
# Join to get species and range info
props_index_MVS <- indices_MVS |>
  group_by(ID) |>
  summarise(
    total = sum(count, na.rm = TRUE),
    very_high = sum(count[riskcat == "Very High"], na.rm = TRUE),
    high = sum(count[riskcat == "High"], na.rm = TRUE),
    perc_high = (high / total) * 100,
    perc_very_high = (very_high / total) * 100,
    perc_both = perc_high + perc_very_high
    )
#####



# Join to get species and range info in one table

# DEPRECATED
# stats_MVG <- props_MVG |>
#   left_join(st_drop_geometry(mammals_sf)[, c("ID", "LISTED_TAXON_ID", "PRESENCE_RANK", "VERNACULAR_NAME","THREATENED_STATUS","TAXON_FAMILY")], by = "ID")

# DEPRECATED
# write_csv(stats_MVG, "Output/data/mamm_stats_MVG.csv", col_name = TRUE)

stats_MVS <- props_MVS |>
  left_join(st_drop_geometry(mammals_sf)[, c("ID", "LISTED_TAXON_ID", "PRESENCE_RANK", "VERNACULAR_NAME","THREATENED_STATUS","TAXON_FAMILY")], by = "ID")

write_csv(stats_MVS, "Output/data/mamm_stats_MVS.csv", col_name = TRUE)

stats_MVS <- read_csv("Output/data/mamm_stats_MVS.csv")

# drop Nas -> habitat not within bounds (e.g. christmas island)
# Drop possible hbitata also
stats_MVS_noNA <- stats_MVS |>
  filter(is.finite(weighted_mean)) |>
  filter(PRESENCE_RANK != 1) # keeps only mammals likely habitat

# This one is for the index (ONLY MVS)
mam_classified_stats <- props_index_MVS |>
  left_join(st_drop_geometry(mammals_sf)[, c("ID", "LISTED_TAXON_ID", "PRESENCE_RANK", "VERNACULAR_NAME","THREATENED_STATUS","TAXON_FAMILY")], by = "ID")

write_csv(mam_classified_stats, "Output/data/mamm_classified_stats_MVS.csv", col_name = TRUE)

mam_classified_stats <- read_csv("Output/data/mamm_classified_stats_MVS.csv")

# cleaning up Nas as above removing chrissy island also
mam_classified_noNA <- mam_classified_stats  |>
  filter(total != 1) |>
  filter(PRESENCE_RANK != 1) # keeps only mammals likely habitat
```

## Mammal trait data

### Trait analyses

```{r}
mam_col <- as.data.frame(mammals_sf) |>
  dplyr::select(LISTED_TAXON_ID, SCIENTIFIC_NAME, VERNACULAR_NAME) |>
  distinct()
  


write_excel_csv(mam_col, "data/processed/mam_names.csv", col_names = TRUE)

COMB_traits <- read_csv("data/raw/ecy3344-sup-0001-datas1/COMBINE_archives/trait_data_imputed.csv")
# Clean up combine traits for necessary info
COMB_traits_small <- COMB_traits |>
  dplyr::select(iucn2020_binomial, phylacine_binomial, fossoriality, dphy_invertebrate, dphy_vertebrate, dphy_plant, det_fruit, det_nect, det_seed, det_plantother, det_diet_breadth_n, trophic_level, foraging_stratum)


# Find species in mammals_sf with no match in COMB_traits
unmatched_species <- mammals_sf |>
  anti_join(COMB_traits_small, by = c("BINOMIAL_NAME" = "iucn2020_binomial"))
mammals_sf <- mammals_sf |>
  mutate(BINOMIAL_NAME= word(SCIENTIFIC_NAME, 1, 2))

# Check which names are not found on the combine dataset
unmatched_names <- unique(unmatched_species$SCIENTIFIC_NAME)
unmatched_names

# Make a tribble of the names of species in the aus dataset which have yet to be updated taxonomically (Notmacropus parma, Osphranter robustus isabellinus, petauroides minor), are unrecognised subspecies on Combine (pteropus melanotus natalis), is sympatric to another species (Rhinolophus reobertsi) or is a binomialisation of a trinomial name (Sminthopsis fuliginosa aitkeni). This allows for them to be selected later. "Antechinus arktos","Antechinus argentus" and "Miniopterus orianae bassanii" simply do not exist on combine in any form. 
synonyms <- tribble(
  ~SCIENTIFIC_NAME,                 ~special_binomial,          ~special_source,
  "Sminthopsis fuliginosa aitkeni", "Sminthopsis aitkeni",      "phylacine_binomial",
  "Rhinolophus robertsi",           "Rhinolophus philippensis", NA,
  "Pteropus natalis",               "Pteropus melanotus",       NA, # species-level
  "Petauroides minor",              "Petauroides volans",       NA,
  "Notamacropus parma",             "Macropus parma",           NA,
  "Osphranter robustus isabellinus","Macropus robustus",        NA,
)

# Accounting for differences in taxonomy
mammals_sf_adjusted <- mammals_sf |>
  left_join(synonyms, by = "SCIENTIFIC_NAME") |>
  mutate(
    BINOMIAL = word(SCIENTIFIC_NAME, 1, 2),      # default binomial
    SOURCE   = if_else(!is.na(special_source), 
                       special_source, "iucn2020_binomial"),
    BINOMIAL = coalesce(special_binomial, BINOMIAL)  # replace with special name if exists
  )

# Split into "normal" vs "special" cases
normal_cases <- mammals_sf_adjusted |> filter(SOURCE == "iucn2020_binomial")
special_cases <- mammals_sf_adjusted |> filter(SOURCE == "phylacine_binomial")

# Join separately
normal_join <- normal_cases |>
  left_join(COMB_traits_small, by = c("BINOMIAL" = "iucn2020_binomial"))

special_join <- special_cases |>
  left_join(COMB_traits_small, by = c("BINOMIAL" = "phylacine_binomial"))

# Combine back together
merged_df <- bind_rows(normal_join, special_join)

# Remove unnecessary columns for analysis
Aus_mams_traits <- st_drop_geometry(merged_df) |>
  filter(PRESENCE_RANK != 1) |>
  dplyr::select(-MAP_TAXON_ID, -MIGRATORY_STATUS, -MARINE, -CETACEAN, -PRESENCE_CATEGORY, -PRESENCE_RANK, -EXTRACT_DATE, -TAXON_GROUP, -TAXON_ORDER, -TAXON_CLASS, -TAXON_PHYLUM, -TAXON_KINGDOM, -OTHER_IDS, -CELL_SIZE, -REGIONS, -ATTRIBUTION, -SPRAT_PROFILE, -SHAPE_Length, -SHAPE_Area, -BINOMIAL_NAME, -BINOMIAL, -special_binomial, -special_source, -SOURCE, -phylacine_binomial, -iucn2020_binomial)



write_excel_csv(as.data.frame(Aus_mams_traits), file = "data/processed/Aus_mam_traits.csv")

# Manual edits done in excel
# Citation for brush tail rabbit rat
# Changes were made to how much of each component makes up its diet
# Firth Ronald S. C., Jefferys Elizabeth, Woinarski John C. Z., Noske Richard A. (2005) The diet of the brush-tailed rabbit-rat (Conilurus penicillatus) from the monsoonal tropics of the Northern Territory, Australia. Wildlife Research 32, 517-523.

# Citations for antechinus arktos

# Habitat:
# Riordan CE, Pearce C, McDonald BJF, Gynther I, Baker AM. Vegetation structure and ground cover attributes describe the occurrence of a newly discovered carnivorous marsupial on the Tweed Shield Volcano caldera, the endangered black-tailed dusky antechinus (Antechinus arktos). Ecol Evol. 2020; 10: 2104–2121. https://doi.org/10.1002/ece3.6045
# Diet:
#Baker, A.M., Mutton, T.Y., Hines, H.B. & Dyck, S.V. (2014) The Black-tailed Antechinus, Antechinus arktos sp. nov.: a new species of carnivorous marsupial from montane regions of the Tweed Volcano caldera, eastern Australia Zootaxa, 3765 (2), 101–133. https://doi.org/10.11646/zootaxa.3765.2.1

# Citations for antechinus argentus
# Habitat:
# Plant diversity and structure describe the presence of a new, threatened Australian marsupial within its highly restricted, post-fire habitat Mason ED, Firn J, Hines HB, Baker AM (2017) Plant diversity and structure describe the presence of a new, threatened Australian marsupial within its highly restricted, post-fire habitat. PLOS ONE 12(8): e0182319. https://doi.org/10.1371/journal.pone.0182319
# Diet:
# Mason Eugene D., Burwell Chris J., Baker Andrew M. (2015) Prey of the silver-headed antechinus (Antechinus argentus), a new species of Australian dasyurid marsupial. Australian Mammalogy 37, 164-169. https://doi.org/10.1071/AM14036

# Citation for yell-bellied glider and mahogany glider diet edit:
# The biology of Australian possums and gliders / edited by Ross L. Goldingay and Stephen M. Jackson (2004) Surrey Beatty & Sons https://catalogue.nla.gov.au/catalog/3257384

Aus_mam_traits_edited <- read_csv("data/processed/Aus_mam_traits_edited.csv")

#Summary info for results
# Filter out species outside study
mammal_filtered <- Aus_mam_traits_edited |>
  filter(!SCIENTIFIC_NAME %in% c("Pteropus natalis", "Crocidura trichura"))

# Summary counts
summary_counts <- mammal_filtered |>
  summarise(
    n_species_total = n(),
    n_families = n_distinct(TAXON_FAMILY),
    n_arborial_foragers = sum(foraging_stratum == "Ar", na.rm = TRUE),
    n_plant_dominant = sum(dphy_plant >= 75, na.rm = TRUE)
  )

# Species count by threatened status
threat_counts <- mammal_filtered |>
  count(THREATENED_STATUS, name = "n_species") |>
  arrange(desc(n_species))

# View results
summary_counts
threat_counts

# now for the scoring (just an example idea)

Aus_mam_traits_scored <- Aus_mam_traits_edited |>
  mutate(
    score = 0 +
      if_else(foraging_stratum == "Ar", 1, 0) +
      if_else(dphy_plant >= 75, 1, 0)
    )

#Add scores to a new spatial object ready for priotizr 
# THis new dataset includes only mammals in the "likely habitat" presence category
mammals_scored_geo <- mammals_sf |>
  filter(PRESENCE_RANK != 1) |>
  dplyr::select(LISTED_TAXON_ID, SCIENTIFIC_NAME, THREATENED_STATUS) |>
  left_join(
    dplyr::select(Aus_mam_traits_scored, SCIENTIFIC_NAME, score), 
    by = "SCIENTIFIC_NAME"
  )

saveRDS(mammals_scored_geo, file = "data/processed/mammals_scored_geo_251003.rds")
mammals_scored_geo <- readRDS("data/processed/mammals_scored_geo_251003.rds")

# Make a df with all the relevant info for each mammal
# Attach scores to the stats tables
stats_MVS_scored <- stats_MVS_noNA |>
  mutate(LISTED_TAXON_ID = as.character(LISTED_TAXON_ID)) |>
  left_join(
    Aus_mam_traits_scored |>
      mutate(LISTED_TAXON_ID = as.character(LISTED_TAXON_ID)) |>
      dplyr::select(LISTED_TAXON_ID, SCIENTIFIC_NAME, score),
    by = "LISTED_TAXON_ID"
  )

mam_class_scored <- mam_classified_noNA |>
  mutate(LISTED_TAXON_ID = as.character(LISTED_TAXON_ID)) |>
  left_join(
    Aus_mam_traits_scored |>
      mutate(LISTED_TAXON_ID = as.character(LISTED_TAXON_ID)) |>
      dplyr::select(LISTED_TAXON_ID, SCIENTIFIC_NAME, score),
    by = "LISTED_TAXON_ID"
  )


mammal_mega_db <- mam_class_scored |>
  left_join(
    stats_MVS_scored |>
      dplyr::select(LISTED_TAXON_ID, nonzero, prop_nonzero, weighted_mean),
    by = "LISTED_TAXON_ID"
  ) |>
  dplyr::select(
    ID,
    LISTED_TAXON_ID, 
    SCIENTIFIC_NAME, 
    TAXON_FAMILY, 
    VERNACULAR_NAME,
    THREATENED_STATUS,
    total, 
    nonzero,
    weighted_mean, 
    high, 
    very_high, 
    prop_nonzero, 
    perc_high, 
    perc_very_high, 
    perc_both, 
    score
  )

write_csv(mammal_mega_db, "Output/data/mammal_mega_db.csv")
mammal_mega_db <- read_csv("Output/data/mammal_mega_db.csv")

# quick check of scores
score_counts <- mammal_mega_db |>
  count(score, name = "n_species") |>
  arrange(score)

score_counts
```

### Risky tables

```{r}

# top 20 by proportion in high risk regions
top20_highrisk <- mammal_mega_db |>
  arrange(desc(perc_both)) |>
  slice_head(n = 20) |>
  dplyr::select(SCIENTIFIC_NAME, perc_both, score) |>
  mutate(
    perc_both = round(perc_both, 2)
    ) |>
  dplyr::mutate(
    SCIENTIFIC_NAME = str_wrap(SCIENTIFIC_NAME, width = 26)  # wrap at 20 chars
  )

# Flextable formatting
ft_top20_highrisk <- flextable(top20_highrisk) |>
  # Header labels
  set_header_labels(
    SCIENTIFIC_NAME = "Scientific Name",
    perc_both = "Percent high risk (%)",
    score = "Score"
  ) |>
  
  # Header styling
  bold(part = "header") |>
  bg(part = "header", bg = "#D3D3D3") |>   # light gray header
  color(part = "header", color = "black") |>
  
  # Zebra striping
  theme_zebra() |>
  
  # Align numeric cols
  align(j = c("perc_both", "score"), align = "center") |>
  
  # Italicise scientific names
  italic(j = "SCIENTIFIC_NAME") |>
  
  # Adjust font size and autofit
  fontsize(size = 10, part = "all") |>
  autofit() |>
  
  # Add black horizontal lines
  hline_top(part = "header", border = fp_border(color="black", width = 2)) |>
  hline_bottom(part = "header", border = fp_border(color="black", width = 2)) |>
  hline(i = 1:nrow(top20_highrisk), border = fp_border(color="black", width = 1))

ft_top20_highrisk

print(my_doc, target = "Output/plots/top20highrisk.docx")

save_as_image(ft_top20_highrisk, path = "Output/images/top20_highrisk_table.png", zoom = 2)

# top 20 by weighted mean regions
top20_weightedmean <- mammal_mega_db |>
  arrange(desc(weighted_mean)) |>
  slice_head(n = 20) |>
  dplyr::select(SCIENTIFIC_NAME, weighted_mean, score) |>
  mutate(
    weighted_mean = round(weighted_mean, 2)
    ) |>
  dplyr::mutate(
    SCIENTIFIC_NAME = str_wrap(SCIENTIFIC_NAME, width = 26)  # wrap at 20 chars
  )

# Flextable formatting
ft_top20_weightedmean <- flextable(top20_weightedmean) |>
  # Header labels
  set_header_labels(
    SCIENTIFIC_NAME = "Scientific Name",
    weighted_mean = "Weighted Mean",
    score = "Score"
  ) |>
  
  # Header styling
  bold(part = "header") |>
  bg(part = "header", bg = "#D3D3D3") |>   # light gray header
  color(part = "header", color = "black") |>
  
  # Zebra striping
  theme_zebra() |>
  
  # Align numeric cols
  align(j = c("weighted_mean", "score"), align = "center") |>
  
  # Italicise scientific names
  italic(j = "SCIENTIFIC_NAME") |>
  
  # Adjust font size and autofit
  fontsize(size = 10, part = "all") |>
  autofit() |>
  
  # Add black horizontal lines
  hline_top(part = "header", border = fp_border(color="black", width = 2)) |>
  hline_bottom(part = "header", border = fp_border(color="black", width = 2)) |>
  hline(i = 1:nrow(top20_weightedmean), border = fp_border(color="black", width = 1))

ft_top20_weightedmean

save_as_image(ft_top20_weightedmean, path = "Output/images/top20_weightedmean_table.png", zoom = 2)


# score 2 by weighted mean regions
top_weightedmean_highscore <- mammal_mega_db |>
  filter(score == 2) |>
  arrange(desc(weighted_mean)) |>
  dplyr::select(SCIENTIFIC_NAME, weighted_mean) |>
  mutate(
    weighted_mean = round(weighted_mean, 2)
    ) |>
  dplyr::mutate(
    SCIENTIFIC_NAME = str_wrap(SCIENTIFIC_NAME, width = 26)  # wrap at 20 chars
  )

# Flextable formatting
ft_top_weightedmean_highscore <- flextable(top_weightedmean_highscore) |>
  # Header labels
  set_header_labels(
    SCIENTIFIC_NAME = "Scientific Name",
    weighted_mean = "Weighted Mean"
  ) |>
  
  # Header styling
  bold(part = "header") |>
  bg(part = "header", bg = "#D3D3D3") |>   # light gray header
  color(part = "header", color = "black") |>
  
  # Zebra striping
  theme_zebra() |>
  
  # Align numeric cols
  align(j = c("weighted_mean"), align = "center") |>
  
  # Italicise scientific names
  italic(j = "SCIENTIFIC_NAME") |>
  
  # Adjust font size and autofit
  fontsize(size = 10, part = "all") |>
  autofit() |>
  
  # Add black horizontal lines
  hline_top(part = "header", border = fp_border(color="black", width = 2)) |>
  hline_bottom(part = "header", border = fp_border(color="black", width = 2)) |>
  hline(i = 1:nrow(top_weightedmean_highscore), border = fp_border(color="black", width = 1))

ft_top_weightedmean_highscore

save_as_image(ft_top20_weightedmean, path = "Output/images/Highscore_weightedmean_table.png", zoom = 2)


# score 2 by weighted mean regions
top_highrisk_highscore <- mammal_mega_db |>
  filter(score == 2) |>
  arrange(desc(perc_both)) |>
  dplyr::select(SCIENTIFIC_NAME, perc_both) |>
  mutate(
    perc_both = round(perc_both, 2)
    ) |>
  dplyr::mutate(
    SCIENTIFIC_NAME = str_wrap(SCIENTIFIC_NAME, width = 26)  # wrap at 20 chars
  )
  

# Flextable formatting
ft_top_highrisk_highscore <- flextable(top_highrisk_highscore) |>
  # Header labels
  set_header_labels(
    SCIENTIFIC_NAME = "Scientific Name",
    perc_both = "Percent high risk (%)"
  ) |>
  
  # Header styling
  bold(part = "header") |>
  bg(part = "header", bg = "#D3D3D3") |>   # light gray header
  color(part = "header", color = "black") |>
  
  # Zebra striping
  theme_zebra() |>
  
  # Align numeric cols
  align(j = c("perc_both"), align = "center") |>
  
  # Italicise scientific names
  italic(j = "SCIENTIFIC_NAME") |>
  
  # Adjust font size and autofit
  fontsize(size = 10, part = "all") |>
  autofit() |>
  
  # Add black horizontal lines
  hline_top(part = "header", border = fp_border(color="black", width = 2)) |>
  hline_bottom(part = "header", border = fp_border(color="black", width = 2)) |>
  hline(i = 1:nrow(top_highrisk_highscore), border = fp_border(color="black", width = 1))

ft_top_highrisk_highscore

save_as_image(ft_top_highrisk_highscore, path = "Output/images/highscore_highrisk_table.png", zoom = 2)

```

### Risk matrix

```{r}
# Make a new dataframe to chuck species into based on their exposure level
heat_df <- mammal_mega_db |>
  mutate(exposure_class = case_when(
    perc_both >= 50 ~ "High exposure",
    perc_both >= 20 ~ "Moderate exposure",
    perc_both > 0   ~ "Low exposure",
    TRUE ~ "No exposure"
  ))
# Summarise each to find how many within each category
heat_test <- heat_df |>
    # Define 3-level risk categorisation
  mutate(
    risk_level = case_when(
      exposure_class == "No exposure" ~ "Low",
      score == 2 & exposure_class %in% c("High exposure", "Moderate exposure") ~ "High",
      score == 2 & exposure_class == "Low exposure" ~ "Medium",
      score == 1 & exposure_class == "High exposure" ~ "High",
      score == 1 & exposure_class == "Moderate exposure" ~ "Medium",
      score == 1 & exposure_class == "Low exposure" ~ "Medium",
      score == 0 & exposure_class == "High exposure" ~ "Medium",
      TRUE ~ "Low"
    ),

    risk_level = factor(risk_level, levels = c("Low", "Medium", "High"))
  )


# Summarise each to find how many within each category
heat_summary <- heat_df |>
  group_by(score, exposure_class) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(
    exposure_class = factor(exposure_class,
      levels = c("No exposure", "Low exposure", "Moderate exposure", "High exposure")
    ),
    score = factor(score, levels = c(0, 1, 2))) |>
  
  # Create all possible combinations
  complete(score, exposure_class, fill = list(n = 0)) |>

    # Define 3-level risk categorisation
  mutate(
    risk_level = case_when(
      exposure_class == "No exposure" ~ "Low",
      score == 2 & exposure_class %in% c("High exposure", "Moderate exposure") ~ "High",
      score == 2 & exposure_class == "Low exposure" ~ "Medium",
      score == 1 & exposure_class == "High exposure" ~ "High",
      score == 1 & exposure_class == "Moderate exposure" ~ "Medium",
      score == 1 & exposure_class == "Low exposure" ~ "Medium",
      score == 0 & exposure_class == "High exposure" ~ "Medium",
      TRUE ~ "Low"
    ),

    risk_level = factor(risk_level, levels = c("Low", "Medium", "High"))
  )

# 3-colour palette (green → yellow → red)
risk_cols <- c(
  "Low" = "#2ECC71",     # green
  "Medium" = "#F1C40F",  # yellow
  "High" = "#E74C3C"     # red
)

riskmatrix <- ggplot(heat_summary, aes(x = exposure_class, y = score, fill = risk_level)) +
  geom_tile(color = "white", linewidth = 0.8) +
  geom_text(aes(label = n), colour = "black", fontface = "bold", size = 7) +
  scale_fill_manual(values = risk_cols, name = "Risk level") +
  labs(
    x = "Exposure class",
    y = "Vulnerability score"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    axis.title = element_text(size = 16, face = "bold"),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14),
  )

ggsave("Output/plots/riskmatrix.png", plot = riskmatrix, width = 10, height = 7, dpi = 300)
```

### Sensitivity tests

```{r}
# Define a function that recalculates and summarises for given thresholds
sensitivity_run <- function(high_cut = 50, moderate_cut = 20, low_cut = 0) {
  
  mammal_mega_db |>
    mutate(exposure_class = case_when(
      perc_both >= high_cut ~ "High exposure",
      perc_both >= moderate_cut ~ "Moderate exposure",
      perc_both > low_cut ~ "Low exposure",
      TRUE ~ "No exposure"
    )) |>
    group_by(score, exposure_class) |>
    summarise(n = n(), .groups = "drop") |>
    mutate(
      exposure_class = factor(exposure_class,
        levels = c("No exposure", "Low exposure", "Moderate exposure", "High exposure")
      ),
      score = factor(score, levels = c(0, 1, 2))
    ) |>
    complete(score, exposure_class, fill = list(n = 0)) |>
    mutate(
      risk_level = case_when(
        exposure_class == "No exposure" ~ "Low",
        score == 2 & exposure_class %in% c("High exposure", "Moderate exposure") ~ "High",
        score == 2 & exposure_class == "Low exposure" ~ "Medium",
        score == 1 & exposure_class == "High exposure" ~ "High",
        score == 1 & exposure_class %in% c("Moderate exposure", "Low exposure") ~ "Medium",
        score == 0 & exposure_class == "High exposure" ~ "Medium",
        TRUE ~ "Low"
      ),
      risk_level = factor(risk_level, levels = c("Low", "Medium", "High")),
      high_cut = high_cut,
      moderate_cut = moderate_cut,
      low_cut = low_cut
    )
}

threshold_grid <- expand.grid(
  high_cut = c(40, 50, 60),
  moderate_cut = c(10, 20, 30),
  low_cut = c(0, 5, 10)
)

sensitivity_results <- purrr::pmap_dfr(threshold_grid, sensitivity_run)

risk_summary <- sensitivity_results|>
  group_by(high_cut, moderate_cut, low_cut, risk_level)|>
  summarise(total_n = sum(n), .groups = "drop")


# Extract unique facet values
low_labels  <- paste0("Low exposure: ", sort(unique(risk_summary$low_cut)), "%")
names(low_labels) <- sort(unique(risk_summary$low_cut))

high_labels <- paste0("High exposure: ", sort(unique(risk_summary$high_cut)), "%")
names(high_labels) <- sort(unique(risk_summary$high_cut))


ggplot(risk_summary, aes(x = factor(moderate_cut),
                         y = total_n,
                         fill = risk_level)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(low_cut ~ high_cut, labeller = labeller(low_cut = as_labeller(low_labels),
                                                     high_cut = as_labeller(high_labels))) +
  scale_fill_manual(values = risk_cols,
                    name = "Risk category") +
  labs(
    x = "Moderate threshold (%)",
    y = "Proportion of species",
    title = "Sensitivity of risk classification to exposure thresholds",
    subtitle = "Faceted by low and high exposure cutoffs"
  ) +
  theme_classic()
ggsave("Output/plots/mam_risk_sensitivity.tiff", width = 10, height = 8, dpi = 300)

# --- numeric sensitivity table --- 
# Define your baseline thresholds
baseline_high <- 50
baseline_mod  <- 20
baseline_low  <- 0

# Extract baseline subset
baseline <- sensitivity_results|>
  filter(high_cut == baseline_high,
         moderate_cut == baseline_mod,
         low_cut == baseline_low)|>
  dplyr::select(score, exposure_class, n_baseline = n)

# Compare each threshold combo against baseline
sensitivity_table <- sensitivity_results|>
  left_join(baseline, by = c("score", "exposure_class"))|>
  mutate(
    n_diff = n - n_baseline,
    pct_diff = 100 * (n - n_baseline) / pmax(1, n_baseline)  # avoid divide-by-zero
  )|>
  group_by(high_cut, moderate_cut, low_cut)|>
  summarise(
    total_abs_change = sum(abs(n_diff), na.rm = TRUE),
    mean_pct_change  = mean(abs(pct_diff), na.rm = TRUE),
    .groups = "drop"
  )|>
  arrange(desc(mean_pct_change))

print(sensitivity_table)


```

#### Table with changes

```{r}
sensitivity_run <- function(high_cut = 50, moderate_cut = 20, low_cut = 0) {
  mammal_mega_db |>
    mutate(exposure_class = case_when(
      perc_both >= high_cut ~ "High exposure",
      perc_both >= moderate_cut ~ "Moderate exposure",
      perc_both > low_cut ~ "Low exposure",
      TRUE ~ "No exposure"
    )) |>
    mutate(
      risk_level = case_when(
        exposure_class == "No exposure" ~ "Low",
        score == 2 & exposure_class %in% c("High exposure", "Moderate exposure") ~ "High",
        score == 2 & exposure_class == "Low exposure" ~ "Medium",
        score == 1 & exposure_class == "High exposure" ~ "High",
        score == 1 & exposure_class %in% c("Moderate exposure", "Low exposure") ~ "Medium",
        score == 0 & exposure_class == "High exposure" ~ "Medium",
        TRUE ~ "Low"
      ),
      high_cut = high_cut,
      moderate_cut = moderate_cut,
      low_cut = low_cut
    )
}

sensitivity_species <- purrr::pmap_dfr(threshold_grid, sensitivity_run)


# --- Baseline combo ---
baseline_combo <- list(high_cut = 50, moderate_cut = 20, low_cut = 0)

# Extract baseline risk level for each species
baseline_df <- sensitivity_species |>
  filter(high_cut == baseline_combo$high_cut,
         moderate_cut == baseline_combo$moderate_cut,
         low_cut == baseline_combo$low_cut) |>
  dplyr::select(SCIENTIFIC_NAME, baseline_risk = risk_level)

# Join baseline with all runs
species_risk_compare <- sensitivity_species |>
  left_join(baseline_df, by = "SCIENTIFIC_NAME") |>
  mutate(
    movement = case_when(
      baseline_risk != "High" & risk_level == "High" ~ "Moved INTO High",
      baseline_risk == "High" & risk_level != "High" ~ "Moved OUT of High",
      TRUE ~ "No change"
    )
  )

# --- Filter only those that changed ---
species_high_modchange <- species_risk_compare |>
  filter(movement != "No change") |>
  group_by(SCIENTIFIC_NAME, moderate_cut, movement) |>
  summarise(
    baseline_risk = first(baseline_risk),
    new_risk = first(risk_level),
    .groups = "drop"
  ) |>
  arrange(moderate_cut)

species_high_modchange

# Create a formatted flextable
ft_species_riskmove <- flextable(species_high_modchange) |>
  # Header labels
  set_header_labels(
    SCIENTIFIC_NAME = "Scientific Name",
    moderate_cut = "Moderate threshold (%)",
    movement = "Movement",
    baseline_risk = "Original risk",
    new_risk = "New risk"
  ) |>
  
  # Header styling
  bold(part = "header") |>
  bg(part = "header", bg = "#D3D3D3") |>   # light gray header
  color(part = "header", color = "black") |>
  
  # Zebra striping
  theme_zebra() |>
  
  # Align numeric cols
  align(j = "moderate_cut", align = "center") |>
  align(j = "movement", align = "center") |>
  
  # Italicise scientific names
  italic(j = "SCIENTIFIC_NAME") |>
  
  # Adjust font size and autofit
  fontsize(size = 10, part = "all") |>
  autofit() |>
  
  # Add black horizontal lines
  hline_top(part = "header", border = fp_border(color="black", width = 2)) |>
  hline_bottom(part = "header", border = fp_border(color="black", width = 2)) |>
  hline(i = 1:nrow(species_high_modchange), border = fp_border(color="black", width = 1))

ft_species_riskmove

# Export to Word
doc <- read_docx() |>
  body_add_flextable(ft_species_riskmove)

print(doc, target = "Output/plots/species_highrisk_movement.docx")
```

### Box plot of families

```{r}
vals_family <- vals_list |>
  left_join(
    stats_MVS_noNA |> 
      dplyr::select(ID, TAXON_FAMILY),
    by = "ID") |>
  filter(!is.na(TAXON_FAMILY))



fambox <- ggplot(vals_family, aes(x = Mrust_val_MVS, y = fct_reorder(TAXON_FAMILY, Mrust_val_MVS, .fun = median))) +
  geom_boxplot(
    colour = "black",
    alpha = 0.7,
    fill = "lightyellow",
    outlier.shape = NA,
    outlier.fill = "red",
    outlier.size = 2
  ) +
  xlim(0, 45) +
  labs(
    x = "Relative myrtle rust hazard intensity",
    y = "Family"
  ) +
  theme(
    axis.text.y = element_text(size = 10, face = "italic"),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  )

ggsave("Output/plots/famil_rust_boxplot.png", plot = fambox, width = 9, height = 7, dpi = 300)
```

### Box plot of score 2

```{r}

# Join exposure values to species info and filter for score = 2
vals_joined <- vals_list |>
  left_join(
    mammal_mega_db |> 
      dplyr::select(ID, SCIENTIFIC_NAME, score, THREATENED_STATUS),
    by = "ID"
  ) |>
  filter(score == 2, !is.na(Mrust_val_MVS))

vals_species_name <- vals_joined |>
  mutate(
    sci_label = case_when(
      SCIENTIFIC_NAME == "Petaurus australis Wet Tropics subspecies" ~ "<i>Petaurus australis</i> (Wet Tropics)",
      SCIENTIFIC_NAME == "Phascolarctos cinereus (combined populations of Qld, NSW and the ACT)" ~ "<i>Phascolarctos cinereus</i>^*^",
      TRUE ~ paste0("<i>", SCIENTIFIC_NAME, "</i>")
    )
  )


# Plot: one box per species
 
score2box <- ggplot(vals_species_name, aes(
  x = reorder(str_wrap(sci_label, 20), Mrust_val_MVS, FUN = median),
  y = Mrust_val_MVS,
  fill = THREATENED_STATUS
)) +
  geom_boxplot(color = "black", outlier.shape = NA, width = 0.7) +
  coord_flip() +
  scale_fill_manual(
    values = c(
      "Critically Endangered" = "#d73025",
      "Endangered" = "#fc8d59",
      "Vulnerable" = "#fee090"
    ),
    name = "Threatened status"
  ) +
  ylim(0, 76) +
  theme(
    axis.text.y = element_markdown(size = 10),
    axis.text.x = element_text(size = 10),
    axis.title = element_text(size = 12, face = "bold"),
    legend.position = "right",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    panel.grid.major.y = element_blank()
  ) +
  labs(
    x = "Scientific name",
    y = "Relative myrtle rust intensity"
  )


ggsave("Output/plots/score2_rust_boxplot.png", plot = score2box, width = 9, height = 7, dpi = 300)
```

# PrioritizR

The purpose of using prioritizr is for finding areas of mammal habitat which are highly important for their conservation.

## Data Preparation

### Mammal layer rasterisation

```{r}



pu <- sxs_thresh_scaled

# Rasterise each mammal feature
mammal_range <- list()
mammal_layers <- list() # removed
mammal_names  <- c()
skipped_ids <- c()
scores <- c()

# for loop - rasterising mammal layer
for (i in seq_len(nrow(mammals_scored_geo))) {
  
  print(paste("Starting iteration:", i))
  
  m <- mammals_scored_geo[i, ]
  
  # Rasterise mammal range (presence = 1, else 0)
  r <- rasterize(vect(m), pu, field = 1, background = 0)
  
  # Intervention: focus on rust hotspots
  rxr <- r * pu # REDACTED
  
  # Check if layer has any non-zero values
  max_val <- suppressWarnings(global(rxr, "max", na.rm = TRUE)[1,1])
  has_nonzero <- !is.na(max_val) && max_val > 0
  
  if (has_nonzero) {
    # Construct name
    nm <- paste0(m$LISTED_TAXON_ID, "_",
                 gsub(" ", "_", m$SCIENTIFIC_NAME))
    
    # Add to lists
    mammal_range[[length(mammal_layers) + 1]]  <- r
    mammal_layers[[length(mammal_layers) + 1]] <- rxr
    mammal_names[length(mammal_names) + 1]     <- nm
    scores[length(scores) + 1]                 <- m$score
    
    cat("✔ Added layer: ", nm)
  } else {
    skipped_ids <- c(skipped_ids, m$LISTED_TAXON_ID)
    cat("⚠ Skipped: ", m$SCIENTIFIC_NAME, " (", m$LISTED_TAXON_ID, ")")
  }
}
# --- Summary report ---
cat("\n===== SUMMARY =====\n")
cat("✔", length(mammal_layers), "species added\n")
cat("⚠", length(skipped_ids), "species skipped\n")
if (length(skipped_ids) > 0) {
  cat("Skipped IDs:", paste(skipped_ids, collapse = ", "), "\n")
}

saveRDS(scores, file = "data/processed/score_list.rds")
scores <- readRDS("data/processed/score_list.rds")

# write each mammal range layer to disk to prevent crash
files <- vapply(seq_along(mammal_range), function(i) {
  f <- file.path("data/processed/mam_rast_layers", paste0("layer_", i, ".tif"))
  writeRaster(mammal_range[[i]], f, overwrite = TRUE)
  f
}, character(1))

# Write each mammal x rust layer to disk
files2 <- vapply(seq_along(mammal_layers), function(i) {
  f2 <- file.path("data/processed/mamxrust_rast_layers", paste0("layer_", i, ".tif"))
  writeRaster(mammal_layers[[i]], f2, overwrite = TRUE)
  f2
}, character(1))


# load in files
files <- list.files("data/processed/mam_rast_layers", 
                    pattern = "\\.tif$", 
                    full.names = TRUE)

files2 <- list.files("data/processed/mamxrust_rast_layers", 
                    pattern = "\\.tif$", 
                    full.names = TRUE)

# Now store both as SPatRAster stacks
mammal_stack <- rast(files)
names(mammal_stack) <- mammal_names

writeRaster(mammal_stack, "data/processed/prioritizr_mammal_stack_250916.tif", overwrite =TRUE)

mammal_stack <- rast("data/processed/prioritizr_mammal_stack_250916.tif")
masked_mammal_ranges <- mask(mammal_stack, sus_x_suit_mvs_scaled)

# Targets - needs to be a rep per species - 30 percent based on government stuff
t1 <- rep(0.3, nlyr(mammal_stack))

# --- file naming stuff ---
# Function to save an object with its name in the filename
save_with_name <- function(object, suffix = "min_short") {
  current_date <- Sys.Date()
  formatted_date <- format(current_date, "%y%m%d")
  
  obj_name <- deparse(substitute(object))  # get the object name as a string
  dir_path <- "Output/data/Prioritizr/"
  filename <- paste0(dir_path, formatted_date, "_", obj_name, "_", suffix, ".rds")
  
  saveRDS(object, filename)
  message("Saved as: ", filename)
}
```

### Costs and budgeting

```{r}
# ---- Weighted costs----
# total combined mammal range raster
mammals_real <- max(mammal_stack, na.rm = TRUE)
# make 0s in mammal_real na
mammals_real_NA <- ifel(mammals_real == 0, NA, 1)
# multiply with myrtle rust cutoff (NA where there's theoretically no rust) to get weighted cost
norustnocost_raw <- mask(sxs_thresh_scaled, sus_x_suit_mvs_scaled)
# everywhere there is a mammal pop in the the myrtle rust value
weighted_cost <- (mammals_real_NA * norustnocost_raw) * 0.01
names(weighted_cost) <- "cost"

# ---- biased costs ----
# this one appoints areas with rust as being lower cost
biased_cost <- (1 / (1 + weighted_cost)) * 0.1

negative_cost = -weighted_cost
exp(negative_cost)

# ---- uniform cost layer ----
norustnocost_1 <- ifel(sxs_thresh_scaled == 0, 0, 1)
uniform_cost <- (mammals_real_NA * norustnocost_1) * 0.01
names(uniform_cost) <- "cost"

## ----- Budgeting and locking in -----

## Uniform
# sum
uniform_cost_values <- values(uniform_cost)
b_uniform <- sum(uniform_cost_values, na.rm = TRUE) - 1 #Add -1 to ensure budget is less than cost
# Uniform cost budget which is 10% of total costs
b_uniform_perc <- b_uniform * 0.1


# locking in cells
uniform_lockin <- which(uniform_cost_values == 0)
# create a lock in raster 
lockin_rast <- rast(uniform_cost)
values(lockin_rast) <- 0
values(lockin_rast)[uniform_lockin] <- 1   # set locked-in cells
lockin_rast <- mask(lockin_rast, uniform_cost)


## Weighted
# sum
weighted_cost_values <- values(weighted_cost)
b_weighted <- sum(weighted_cost_values, na.rm = TRUE) - mean(weighted_cost_values, na.rm = TRUE)
# locked in cells
weighted_lockin <- which(weighted_cost_values == 0)

## Biased
# sum
biased_cost_values <- values(biased_cost)
b_biased <- sum(biased_cost_values, na.rm = TRUE) - mean(biased_cost_values, na.rm = TRUE)

#locked in cells
biased_locking <- which(biased_cost_values == 0.1)
```

## Prioritisation scenarios

### DEPRECATED - Scenario 1: Uniform Cost + Unranked features -

```{r}
# # ----- SCENARIO 1: Uniform Cost + Unweighted and unranked features -----
# # More complicated objective - works well 
# p1_0.3 <- problem(uniform_cost, mammal_stack, run_checks = TRUE) |>
#   add_min_shortfall_objective(budget = b_uniform) |>
#   add_relative_targets(t1) |>
#   add_locked_in_constraints(lockin_rast) |>
#   #add_feature_weights(mammals_scored_geo$score[1:nlyr(features_masked)]) |> # The weights make a big difference
#   add_binary_decisions() |>
#   add_gurobi_solver(gap = 0.1)
# 
# # Create a filename with the custom formatted date
# save_with_name(p1_0.3)
# 
# # Solve & importance
# s1_0.3 <- solve(p1_0.3)
# summary(s1_0.3) # this is telling me its one everywhere - whacky 
# 
# # Create a filename with the custom formatted date
# save_with_name(s1_0.3)
# 
# 
# # imp <- eval_replacement_importance(p, s) This takes bloody forever
# imp1_0.3 <- eval_rank_importance(p1_0.3, s1_0.3, n=100) # This is quick
# save_with_name(imp1_0.3)
# 
# # Quick plots to check
# plot(s1_0.3, main="Scenario 1: Intervention scenario (focus on rust hotspots)")
# plot(imp1_0.3, main="Scenario 1: Rank importance")
```

### DEPRECATED - Scenario 2: Weighted Cost + Unranked features

`{# {r} # # ----- SCENARIO 2: Weighted Cost + Unweighted and unranked features ----- # # More complicated objective - works well  # p2_0.3 <- problem(weighted_cost, mammal_stack, run_checks = TRUE) |> #   add_min_shortfall_objective(budget = b_weighted) |> #   add_relative_targets(t1) |> #   add_locked_in_constraints(weighted_lockin) |> #   #add_feature_weights(mammals_scored_geo$score[1:nlyr(features_masked)]) |> # The weights make a big difference #   add_binary_decisions() |> #   add_gurobi_solver(gap = 0.1) #  # save_with_name(p2_0.3) #  # # Solve & importance # s2_0.3 <- solve(p2) # summary(s2)  # save_with_name(s2_0.3) #  #  # # imp <- eval_replacement_importance(p, s) This takes bloody forever # imp2_0.3 <- eval_rank_importance(p2, s2, n=100) # This is quick #  #  # # Quick plots to check # plot(s2, main="Scenario 2: Intervention scenario (focus on rust hotspots)") # plot(imp2, main="Scenario 2: Rank importance")`

### Scenario 3: Uniform Cost + Ranked features

```{r}
# ----- SCENARIO 3: Uniform Cost + Weighted features -----
# More complicated objective - works well 
p3_0.3 <- problem(uniform_cost, mammal_stack, run_checks = TRUE) |>
  add_min_shortfall_objective(budget = b_uniform) |>
  add_relative_targets(t1) |>
  add_locked_in_constraints(uniform_lockin) |>
  add_feature_weights(scores) |> # The weights make a big difference
  add_binary_decisions() |>
  add_gurobi_solver(gap = 0.1)

save_with_name(p3_0.3)

# Solve & importance
s3_0.3 <- solve(p3_0.3)
summary(s3_0.3) 
save_with_name(s3_0.3)

# imp <- eval_replacement_importance(p, s) This takes bloody forever
imp3_0.3 <- eval_rank_importance(p3_0.3, s3_0.3, n=100) # This is quick
save_with_name(imp3_0.3)

# Quick plots to check
plot(s3_0.3, main="Scenario 3: Intervention scenario (focus on rust hotspots)")
plot(imp3_0.3, main="Scenario 3: Rank importance")

```

### Scenario 4: Weighted Cost + Ranked features

```{r}
# ----- SCENARIO 4: Weighted Cost + Weighted features -----
# More complicated objective - works well 
p4_0.3 <- problem(weighted_cost, mammal_stack, run_checks = TRUE) |>
  add_min_shortfall_objective(budget = b_weighted) |>
  add_relative_targets(t1) |>
  add_locked_in_constraints(weighted_lockin) |>
  add_feature_weights(scores) |> # The weights make a big difference
  add_binary_decisions() |>
  add_gurobi_solver(gap = 0.1)

save_with_name(p4_0.3)

# Solve & importance
s4_0.3 <- solve(p4_0.3)
summary(s4_0.3) 

save_with_name(s4_0.3)

# imp <- eval_replacement_importance(p, s) This takes bloody forever
imp4_0.3 <- eval_rank_importance(p4_0.3, s4_0.3, n=100) # This is quick
save_with_name(imp4_0.3)

# Quick plots to check
plot(s4_0.3, main="Scenario 4: Intervention scenario (focus on rust hotspots)")
plot(imp4_0.3, main="Scenario 4: Rank importance")
```

### DEPRECATED - Scenario 5: Biased Cost + Unranked features - DEPRECATED

```{r}
# # More complicated objective - works well 
# 
# p5 <- problem(biased_cost, mammal_stack, run_checks = TRUE) |>
#   add_min_shortfall_objective(budget = b_biased) |>
#   add_relative_targets(t1) |>
#   #add_feature_weights(scores) |> # The weights make a big difference
#   add_binary_decisions() |>
#   add_gurobi_solver(gap = 0.1)
# 
# saveRDS(p5, file = "Output/data/Pioritizr/250922_p5_min_short.rds")
# 
# # Solve & importance
# s5 <- solve(p5)
# summary(s5) # this is telling me its one everywhere - whacky 
# saveRDS(s5, file = "Output/data/Pioritizr/250922_s5_min_short.rds")
# 
# 
# # imp <- eval_replacement_importance(p, s) This takes bloody forever
# imp5 <- eval_rank_importance(p5, s5, n=100) # This is quick
# saveRDS(imp5, file = "Output/data/Pioritizr/250922_imp5_min_short.rds")
# 
# # Quick plots to check
# plot(s5, main="Scenario 5: Intervention scenario (focus on rust hotspots)")
# plot(imp5, main="Scenario 5: Rank importance")
```

### Scenario 6: Biased Cost + Ranked features

```{r}
p6_0.3 <- problem(biased_cost, mammal_stack, run_checks = TRUE) |>
  add_min_shortfall_objective(budget = b_biased) |>
  add_relative_targets(t1) |>
  add_locked_in_constraints(biased_locking) |>
  add_feature_weights(scores) |> # The weights make a big difference
  add_binary_decisions() |>
  add_gurobi_solver(gap = 0.1)

save_with_name(p6_0.3)

# Solve & importance
s6_0.3 <- solve(p6_0.3)
summary(s6_0.3) # this is telling me its one everywhere - whacky 

save_with_name(s6_0.3)

# imp <- eval_replacement_importance(p, s) This takes bloody forever
imp6_0.3 <- eval_rank_importance(p6_0.3, s6_0.3, n=100) # This is quick

save_with_name(imp6_0.3)

# Quick plots to check
plot(s6_0.3, main="Scenario 6: Intervention scenario (focus on rust hotspots)")
plot(imp6_0.3, main="Scenario 6: Rank importance")
```

### Scenario 7: Penalty test

```{r}
# More complicated objective - works well 
p_penalty <- problem(uniform_cost, mammal_stack, run_checks = TRUE) |>
  add_min_shortfall_objective(budget = b_uniform) |>
  add_relative_targets(t1) |>
  add_locked_in_constraints(uniform_lockin) |>
  add_feature_weights(scores) |> # The weights make a big difference
  add_locked_in_constraints(uniform_lockin) |>
  add_linear_penalties(penalty = 100, data = negative_cost) |>
  add_binary_decisions() |>
  add_gurobi_solver(gap = 0.1)

#p_penalty <- readRDS("Output/data/Prioritizr/251010_p_penalty_min_short.rds")
save_with_name(p_penalty)

# Solve & importance
s_penalty <- solve(p_penalty)
summary(s_penalty) 

save_with_name(s_penalty)

# imp <- eval_replacement_importance(p, s) This takes bloody forever
imp_penalty <- eval_rank_importance(p_penalty, s_penalty, n=100) # This is quick


save_with_name(imp_penalty)

plot(imp_penalty)


```

### Visualisations

```{r, warning = FALSE}

# Load in all importance rasters from file
imp_penalty <- readRDS("Output/data/Prioritizr/251014_imp_penalty_min_short.rds")
imp6_0.3 <- readRDS("Output/data/Prioritizr/251014_imp6_0.3_min_short.rds")
imp4_0.3 <- readRDS("Output/data/Prioritizr/251014_imp4_0.3_min_short.rds")
imp3_0.3 <- readRDS("Output/data/Prioritizr/251010_imp3_0.3_min_short.rds")

# Load in shapefule
east_states_shp <- st_read("data/processed/east_states_shp.shp")
east_states_shp_wgs84 <- st_transform(east_states_shp, crs(lockin_rast))

# Project all to Albers Equal area EPSG:3577 for plotting
lock_overlay <- project(lockin_rast, sxs_alb, method = "near")
impp_alb <- project(imp_penalty, lock_overlay)
imp6_alb <- project(imp6_0.3, lock_overlay)
imp4_alb <- project(imp4_0.3, lock_overlay)
imp3_alb <- project(imp3_0.3, lock_overlay)


# Crop all to Australian east states shapefile
impp_east <- crop(imp_penalty, east_states_shp_wgs84)
imp6_east <- crop(imp6_0.3, east_states_shp_wgs84)
imp4_east <- crop(imp4_0.3, east_states_shp_wgs84)
imp3_east <- crop(imp3_0.3, east_states_shp_wgs84)
lock_overlay_east <- crop(lockin_rast, east_states_shp_wgs84)

# Mask by states
impp_east <- mask(impp_east, east_states_shp_wgs84)
imp6_east <- mask(imp6_east, east_states_shp_wgs84)
imp4_east <- mask(imp4_east, east_states_shp_wgs84)
imp3_east <- mask(imp3_east, east_states_shp_wgs84)
overlay_east <- mask(lock_overlay_east, east_states_shp_wgs84)

# raster dataframes for GG
# Penalty importance
df_impp <- as.data.frame(impp_east, xy = TRUE, na.rm = TRUE)
# df_impp$x <- round(df_impp$x, 0)
# df_impp$y <- round(df_impp$y, 0)
# Scenario 6 importance
df_imp6 <- as.data.frame(imp6_east, xy = TRUE, na.rm = TRUE)
# df_imp6$x <- round(df_imp6$x, 0)
# df_imp6$y <- round(df_imp6$y, 0)
# Scenario 4 importance
df_imp4 <- as.data.frame(imp4_east, xy = TRUE, na.rm = TRUE)
# df_imp4$x <- round(df_imp4$x, 0)
# df_imp4$y <- round(df_imp4$y, 0)
# Scenario 3 importance
df_imp3 <- as.data.frame(imp3_east, xy = TRUE, na.rm = TRUE)
# df_imp3$x <- round(df_imp3$x, 0)
# df_imp3$y <- round(df_imp3$y, 0)
# Overlay
df_overlay <- as.data.frame(overlay_east, xy = TRUE, na.rm = TRUE)

# Simplify lines on states
states_simplified <- st_simplify(east_states_shp, dTolerance = 1000)
states_simplified <- st_simplify(east_states_shp_wgs84, dTolerance = 1000)

p_impp <- ggplot() +
  geom_sf(data = states_simplified, fill = "white", colour = NA, linewidth = 0, lineend = "round", linejoin = "round") +
  geom_raster(data = df_impp, aes(x, y, fill = rs)) +
  geom_raster(data = subset(df_overlay, cost == 1), aes(x = x, y = y), fill = "grey80", alpha = 1) +
  geom_sf(data = states_simplified, fill = NA, colour = "black", linewidth = 0.15, lineend = "round", linejoin = "round") +
  scale_fill_viridis_c(
    option = "viridis",
    name = "Importance",
    values = scales::rescale(c(0.0, 0.20, 0.50, 0.70, 0.9, 1.00)),
    na.value = "transparent"
  ) +
  coord_sf(expand = TRUE) +
  labs(title = "Uniform cost with penalty") +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

p_impp

p_imp6 <- ggplot() +
  geom_sf(data = states_simplified, fill = "white", colour = NA, linewidth = 0, lineend = "round", linejoin = "round") +
  geom_raster(data = df_imp6, aes(x, y, fill = rs)) +
  geom_raster(data = subset(df_overlay, cost == 1), aes(x = x, y = y), fill = "grey80", alpha = 1) +
  geom_sf(data = states_simplified, fill = NA, colour = "black", linewidth = 0.15, lineend = "round", linejoin = "round") +
  scale_fill_viridis_c(
    option = "viridis",
    name = "Importance",
    values = scales::rescale(c(0.0, 0.20, 0.50, 0.70, 0.9, 1.00)),
    na.value = "transparent"
  ) +
  coord_sf(expand = TRUE) +
  labs(title = "Biased cost") +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

p_imp6

p_imp4 <- ggplot() +
  geom_sf(data = states_simplified, fill = "white", colour = NA, linewidth = 0, lineend = "round", linejoin = "round") +
  geom_raster(data = df_imp4, aes(x, y, fill = rs)) +
  geom_raster(data = subset(df_overlay, cost == 1), aes(x = x, y = y), fill = "grey80", alpha = 1) +
  geom_sf(data = states_simplified, fill = NA, colour = "black", linewidth = 0.15, lineend = "round", linejoin = "round") +
  scale_fill_viridis_c(
    option = "viridis",
    name = "Importance",
    values = scales::rescale(c(0.0, 0.20, 0.50, 0.70, 0.9, 1.00)),
    na.value = "transparent"
  ) +
  coord_sf(expand = TRUE) +
  labs(title = "Rust weighted cost") +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

p_imp4

p_imp3 <- ggplot() +
  geom_sf(data = states_simplified, fill = "white", colour = NA, linewidth = 0, lineend = "round", linejoin = "round") +
  geom_raster(data = df_imp3, aes(x, y, fill = rs)) +
  geom_raster(data = subset(df_overlay, cost == 1), aes(x = x, y = y), fill = "grey80", alpha = 1) +
  geom_sf(data = states_simplified, fill = NA, colour = "black", linewidth = 0.15, lineend = "round", linejoin = "round") +
  scale_fill_viridis_c(
    option = "viridis",
    name = "Importance",
    values = scales::rescale(c(0.0, 0.20, 0.50, 0.70, 0.9, 1.00)),
    na.value = "transparent"
  ) +
  coord_sf(expand = TRUE) +
  labs(title = "Uniform cost") +
  annotation_north_arrow(location = "tr", which_north = "true",
                         style = north_arrow_fancy_orienteering) +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black"),
        panel.grid.major = element_line(color = "grey50"))

p_imp3

combo_plot <- ((p_imp3 + p_impp) / (p_imp4 + p_imp6)) +
   plot_layout(
    guides = "collect"
  ) &
  theme(
    legend.position = "right",
    legend.justification = "center",
    legend.text = element_text(size = 11),
    legend.title = element_text(size = 13),
    legend.key.height = unit(1.3, "cm"),
    legend.key.width = unit(1, "cm"),
    legend.box.spacing = unit(-2.3, "cm"),
    plot.margin = margin(5, 5, 5, 5)
  )


combo_plot
ggsave("Output/plots/combo_importance.tiff",
       combo_plot, width = 10, height = 14, dpi = 300)
```
